---
title: "Marketing report part 2"
date: 1900-01-01
author: 
  - Norah Jones   
  - Bill Gates
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: visual
---

```{r }
#| label: libraries
#| echo: false

# Check out https://quarto.org/docs/output-formats/html-basics.html
## to get started with html documents

# Check out https://quarto.org/docs/reference/formats/html.html
## to examine all the possible options for html documents

# You can control the execution options of each code 
# chunk
## For more information check out 
### https://quarto.org/docs/computations/execution-options.html
# You can include the libraries you are going to use here
library(sweep)
```

# Exploring data, visualizing it and applying statistical tests

## Introduction

Once your group has finish filling all the sections, add some text here describing the principal findings. Be straightforward and concise taking into account that it is a executive report.

## Import data

Add some text if your group consider to be necessary.

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales
```

## Relation between continuous variables

Use the available functions from base R or the `tidyverse` to create a simple scatterplot between `price` and `quantity` colored by `frame`

```{r}
# You can use plot and legend from base R 
# or ggplot, geom_point and labs from the tidyverse
## If you use ggplot, geom_point and labs remember to load the 
## the library in the first code chunk: library(tidyverse)
### For more information about using geom_point and some examples check out
#### https://ggplot2.tidyverse.org/reference/geom_point.html
```

Generate the same scatterplot but applying a logarithmic transformation to `price` and `quantity`

```{r}
# If you use base R you can use the argument
# log ="xy" in plot
# If you use the tidyverse you can use the functions
# scale_x_continuous and scale_y_continuous by using
# the argument transform (nowdays trans is deprecated) where you can try the option "log10"
## Check out https://ggplot2.tidyverse.org/reference/scale_continuous.html for more information
## and some examples
```

## Relation between categorical variables

Use the available functions from the `tidyverse` to create a simple scatterplot between `frame` and `category.secondary`

Also with this kind of plot you can identify the [product gap](https://en.wikipedia.org/wiki/Gap_analysis) in relation to potential products that are not offer

```{r}
# When using the tidyverse use ggplot, geom_point,
# labs
## In the position argument don't forget to use 
## position_jitter
### Check out https://ggplot2.tidyverse.org/reference/position_jitter.html  for more information
## and some examples
```

## Descriptive statistics by group

### Mean revenue by `(category.secondary, frame)`



### Percentage of total revenue generated by each `(category.secondary, frame)`



### Mean revenue by `bikeshop.name`



### Percentage of total revenue generated by each `bikeshop.name`



## Statistical tests

### Testing group frequencies: `chisq.test()`

Pending: Percentage of total revenue

(category.secondary, frame)

bikeshop.name

### Testing group means: `t.test()`

Pending: between frames

### Analysis of Variance (ANOVA)

Model: price ~ category.secondary + frame (base model)