---
title: "Marketing report part 3"
date: 1900-01-01
author: 
  - Norah Jones   
  - Bill Gates
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: visual
---

```{r }
#| label: libraries
#| echo: false

# Check out https://quarto.org/docs/output-formats/html-basics.html
## to get started with html documents

# Check out https://quarto.org/docs/reference/formats/html.html
## to examine all the possible options for html documents

# You can control the execution options of each code 
# chunk
## For more information check out 
### https://quarto.org/docs/computations/execution-options.html
# You can include the libraries you are going to use here
library(sweep)
```

# Modeling

## Introduction

Once your group has finish filling all the sections, add some text here describing the principal findings. Be straightforward and concise taking into account that it is an executive report.

## Import data

Add some text if your group consider to be necessary.

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales
```

## The product gap

Add some text pointing out the principal findings as a result of generating the following visualization about the [product gap](https://en.wikipedia.org/wiki/Gap_analysis):

Use the available functions from the `tidyverse` to create a visualization so the Research and Development (R & D) division can understand the [product gap](https://en.wikipedia.org/wiki/Gap_analysis)

Brake the process in 2 steps:

-   Prepare the data

    -   You can use `select`, `group_by`, `summarise`, `ungroup`, `arrange`, `mutate` and `fct_reorder`

```{r}
# Use bike_sales to calculate the total revenue
# grouped by model, category.secondary and frame

## You can proceed in the following way
### This is only a template you can modify the code
### if you want to add or apply another way to 
### prepare the data
#### Remember to load the tidyverse library: library(tidyverse)

# total_sales_by_model_second_category_frame <- bike_sales |> 
#   select(model, category.secondary, frame, price.ext) |> 
#   group_by(model, category.secondary, frame) |> 
#   summarise(total_revenue = sum(price.ext)) |> 
#   ungroup() |> 
#   arrange(desc(total_revenue)) |> 
#   mutate(category.secondary = fct_reorder(.f = category.secondary,
#                                           .x = total_revenue))
# 
# total_sales_by_model_second_category_frame
```

-   Generate an informative plot

    -   You can use `ggplot`, `geom_point`, `position_jitter`, `scale_y_continuous`, `scales::label_dollar` and `facet_wrap`

    -   You can also use `labs` to personalize the labels of the x-axis, y-axis and title and change the colors if you want

Using `facet_wrap` you can identify which combinations between `category.secondary` and `frame` are empty without data where they correspond to the [product gap](https://en.wikipedia.org/wiki/Gap_analysis)

Also check out the combinations that you reported in the first marketing report to verify that your plot is reflecting this aspect

```{r}
# You can visualize the product gap by generating 
# group visualizations by category.secondary
# when inspecting the relation between frame and
# total revenue

## You can proceed in the following way
### This is only a template you can modify the code
### if you want to add or apply another way to 
### visualize the data or personalize the plot

# total_sales_by_model_second_category_frame |> 
#   ggplot() + 
#   geom_point(aes(x = frame, y = total_revenue), 
#              position = position_jitter(width = 0.2, 
#                                         height = 0.2)) +
#   scale_y_continuous(labels = scales::label_dollar()) +
#   facet_wrap(facets = vars(category.secondary))
```

## Determine the price of a new product

Add some text pointing out the principal findings as a result of applying a linear statistical model to determine the price of a new product

### Selecting a new potential product

Once you have identify the combination of categories that belong to the product gap please choose a product that is not offer by the corporation so the Research and Development (R & D) division can develop it

According to the results of the first marketing report remember that the categories that belong to the product gap are:

-   Cyclocross Aluminum
-   Fat Bike Carbon
-   Over Mountain Aluminum
-   Sport Carbon
-   Triathalon Aluminum

Taking into account that it is difficult for the Research and Development (R & D) division to develop 5 new products please recommend this division a new product to develop by pointing out a sound reason based on you marketing expertise

### Building a model to price a new product

Once you have choose the new product you want to recommend to the Research and Development (R & D) division in this section you are going to determine the price of this new product using similar products already offer by the corporation

#### Choosing a model

Using the results developed in the second marketing report in relation to the **Analysis of Variance (ANOVA)** section choose a model that you consider appropiate.

Recall that you explored the following models, which are organized by level of complexity from the lowest to the highest:

-   `price ~ category.secondary`
-   `price ~ category.secondary + frame`
-   `price ~ category.secondary*frame`
-   `price ~ category.secondary*frame*bikeshop.state`

**Note 1**: If the model is to complex like `price ~ category.secondary*frame*bikeshop.state` you will obtain a considerable parameters but if the model is to simple like `price ~ category.secondary` you will not be able to obtain good predictions. Choose the level of complexity that you consider appropriate to help the Research and Development (R & D) division and to determine the new price of the new product

**Note 2**: If you find that the model generates silly results, like for example negative or 0 values for the price, abandon the choosen model and use another model or change the product gap combination or discard particular locations where the product will going to be offer.

#### Fit the model

One you have choose the model, fit it using the function `lm` and explore the results of the model with the `tidy` function from the `tidymodels` library:

```{r}
# For example if you choose the model
# price ~ category.secondary*frame*bikeshop.state 
# you will fit the following model
## Remember to load the tidymodels library: library(tidymodels) 

# model <- lm(formula = price ~ category.secondary*frame*bikeshop.state,
#             data = bike_sales)
# 
# model |> tidy()
```

#### Predict the price of the new product

Once you fit the model and taking into account the selected new potential product predict the new price for the new product

Remember to create a tibble or data frame with the new data you want to predict

In this part you can use the functions `predict`, `enframe` and `bind_cols` to present your results in an organized way

```{r}
# For example if you choose the model
# price ~ category.secondary*frame*bikeshop.state
# and the new product to develop as Cyclocross Aluminium
# you proceed in the following way

## In this case because the model 
## price ~ category.secondary*frame*bikeshop.state
## include the bikeshop.state you need to select at least 
## one state to make predictions

### Lets predict the price for 3 states and the combination
### Triathalon Aluminum: 
#### "AZ": Arizona 
#### "CA": California
#### "CO": Colorado
#### category.secondary = "Triathalon"
#### frame = "Aluminum"

# new_data <- tibble(category.secondary = c("Triathalon", "Triathalon", "Triathalon"),
#                    frame = c("Aluminum", "Aluminum", "Aluminum"),
#                    bikeshop.state = c("AZ", "CA", "CO"))
# 
# predict(object = model, newdata = new_data) |>
#   enframe(name = "observation", value = "price_prediction") |>
#   bind_cols(new_data)

### If you check this example you will get
#   observation price_prediction category.secondary frame    bikeshop.state
#   <chr>                  <dbl> <chr>              <chr>    <chr>         
# 1 1                     -705.  Triathalon         Aluminum AZ            
# 2 2                       95.8 Triathalon         Aluminum CA            
# 3 3                    -1157.  Triathalon         Aluminum CO

#### In this case it only make sense to offer this product in California (CA)
#### In the remaining states, the product is not worth offering. In such cases 
#### you would have to give it free to the public.
```
