---
title: "Comparing Groups: Tables and Visualizations"
author: 
  - Luis Francisco Gomez
  - Sara Perez
format: 
  html:
    embed-resources: true
    toc: true
    number-sections: true
    code-fold: true
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(skimr)
library(DT)
library(scales)
```

# Import data

```{r}
#| message: false

segmentation <- read_csv(file = "../000_data/005_data_chapter5.csv")
```

# Transform

```{r}
segmentation <- segmentation |> 
  mutate(gender = factor(x = gender,
                         ordered = FALSE),
         kids = as.integer(kids),
         ownHome = factor(x = ownHome,
                          ordered = FALSE),
         subscribe = factor(x = subscribe,
                            ordered = FALSE),
         Segment = factor(x = Segment,
                          ordered = FALSE))
```

# Summarize data

Check out the `income` variable because it has negative values

```{r}
segmentation |> 
  skim()
```

# Summary statistics by groups

Question 1: Are there differences between segments in relation to income?

```{r}
segmentation |> 
  select(Segment, income) |> 
  group_by(Segment) |> 
  summarise(mean_income = mean(income)) |> 
  arrange(desc(mean_income)) |> 
  datatable()
```

Question 2: Are there differences between segments in relation to the number of kids?

```{r}
#| message: false

segmentation |> 
  group_by(ownHome, Segment) |> 
  summarise(number_kids = sum(kids)) |> 
  ungroup() |> 
  arrange(desc(number_kids)) |> 
  datatable()
```

Question 3: What are the percentage of subscribers by Segment and by the status of owning a house?

```{r}
subscriber_by_segment_home_ownership <- segmentation |> 
  count(subscribe, ownHome, Segment) |> 
  group_by(ownHome, Segment) |> 
  mutate(number_percentage = (n / sum(n))*100) |> 
  ungroup() |> 
  arrange(Segment, ownHome)

subscriber_by_segment_home_ownership |> 
  datatable()
```

## Visualization by groups

### More than 2 groups

When grouping by ownership house status and segment the percentage of subscribers of my product is low.

```{r}
subscriber_by_segment_home_ownership |> 
  ggplot() +
  geom_col(aes(x = subscribe, 
               y = number_percentage),
           color = "black",
           fill = "#7ED4AD") +
  facet_wrap(facets = vars(ownHome, 
                           Segment), 
             nrow = 2, ncol = 4)
```

### Proportions by group

Question 4: What proportion of clients are subscribers by segment?

```{r}
prop_table <- segmentation |> 
  select(Segment, subscribe) |> 
  count(Segment, subscribe) |> 
  group_by(Segment) |> 
  mutate(pct_subscribers = n / sum(n)) |> 
  ungroup() |> 
  filter(subscribe == "subYes") |> 
  mutate(Segment = fct_reorder(.f = Segment,
                               .x = pct_subscribers))

prop_table |> 
  ggplot() + 
  geom_col(aes(x = pct_subscribers, 
               y = Segment),
           color = "black",
           fill = "#7ED4AD") + 
  labs(x = "Proportion of subscribers",
       y = "",
       title = "Proportions of subscribers by segment")
```

### Grouping with a continuous variable

```{r}
#| message: false

mean_income_by_seg_ownhome <- segmentation |> 
  select(income, ownHome, Segment) |> 
  group_by(Segment, ownHome) |> 
  summarise(mean_income = mean(income)) |> 
  ungroup()

mean_income_by_seg_ownhome |> 
  ggplot() + 
  geom_col(aes(x = Segment, y = mean_income,
               fill = ownHome),
           position = position_dodge(),
           color = "black") + 
  scale_fill_manual(values = c("#7ED4AD", 
                               "#D76C82")) +
  labs(x = "",
       y = "Mean disposable income (US Dollars)",
       fill = "Owning home status",
       title = "Comparing of mean income by segment and owning home status")
  
```

### Grouping with a continuous variable using more than 1 category

```{r}
income_distribution_by_seg_ownhome <- segmentation |> 
  select(income, ownHome, Segment) |>
  mutate(Segment = fct_reorder(.f = Segment,
                               .x = income))
  
income_distribution_by_seg_ownhome |>    
  ggplot() + 
  geom_boxplot(aes(x = income, y = Segment,
                   fill = ownHome)) + 
  scale_fill_manual(values = c("#7ED4AD", 
                               "#D76C82")) +
  scale_x_continuous(labels = label_dollar(scale = 0.01)) +
  facet_wrap(facets = vars(ownHome)) + 
  labs(x = "Disposable income in thousands (USD dollars)",
       y = "",
       fill = "Owning home status",
       title = "Distribution of income by owning home status and segment")
```
