---
title: "An on overview of the R language"
author: 
  - Luis Francisco Gomez
  - Juan Perez
  - Juana de Heredia
format: 
  html:
    embed-resources: true
    code-fold: true
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(skimr)
library(DT)
library(sweep)
```

# Importing data

This is a data set related to the satisfaction of a group of clients and also its corresponding segment.

```{r}
#| message: false
# satisfaction_data <- read_csv(file = "http://goo.gl/UDv12g")
satisfaction_data <- read_csv(file = "../000_data/002_data_chapter2.csv")
satisfaction_data |> 
  head(n = 5) |> 
  datatable()
```

# Tranforming data

Taking into account that the Segment variable is nominal we need to transform this column in the data.

```{r}
satisfaction_data <-  satisfaction_data |> 
  mutate(Segment = factor(x = Segment,
                          ordered = FALSE))
```

# Summarizing data

```{r}
satisfaction_data |> skim()
```

# Project practice

## Import data

```{r}
bike_sales <- bike_sales
```

## Answering questions

-   Question 1: ¿What are the bike shops that generate more revenue?

```{r}
revenue_by_bikeshop <- bike_sales |> 
  select(customer.id, bikeshop.name,
         quantity, price) |> 
  mutate(revenue = quantity * price) |> 
  group_by(bikeshop.name) |> 
  summarise(total_revenue = sum(revenue))

revenue_by_bikeshop |> 
  datatable()
```

