---
title: "Specific instructions: marketing report part 3"
execute: 
  echo: true
  warning: false
format: 
  html:
    number-sections: true
    toc: true
    embed-resources: true
    css: ../000_css/styles.css
bibliography: ../000_references/marketing_research_and_analytics.bib
editor: visual
---

# Business Case

[Disclaimer:]{style="color: red;"} This business case was taken from the course [Data Science for Business Part 1](https://university.business-science.io/p/ds4b-101-r-business-analysis-r) offered by the company [Business Science](https://www.business-science.io/) and adapted to be in line with the topics covered in [@chapman_r_2019].

You and your team work for a corporation located in Wilton, Connecticut, United States that supplies bicycle frames and other components related to bicycles to different bicycle shops through the United States.

Your team is assigned to complete 2 tasks:

-   Support the Research and Development (R & D) division in identifying potential new products and pricing them by using data collected from the bicycle shops.

-   Support the marketing team in the creation of a marketing segmentation clustering model by using data collected from to the bicycle shops to offer more personalized products and messaging them.

# Packages

You will need to install a set of packages that are going to be used along the course to extend the functionality of R. You can find the packages that are going to be used in 2 locations. Choose the option of your preference:

-   <https://github.com/luifrancgom/marketing_research_and_analytics/> \> 000_packages_for_the_course \> 001_script_install_packages.R

-   Primer Corte 30% \> Learning Resources \> Support material \> Packages for the course \> 001_script_install_packages.R

Run the full script using RStudio IDE to install the packages. If you get the following message: **Do you want to install from source the packages which need compilation?**, just say no and keep going.

-   If you are curious about this message check out this [post](https://forum.posit.co/t/meaning-of-common-message-when-install-a-package-there-are-binary-versions-available-but-the-source-versions-are-later/2431) and the answer by Jenny Bryan

-   If you understand what is the meaning of **built from source locally** and you want to do this just install the following tools:

    -   For **Windows Users**: <https://cran.r-project.org/bin/windows/Rtools/> \> *Choose the RTools version according to your R version* (type `R.version` in the RStudio console if you don't remember the R version you have installed)

    -   For **Mac Users**: <https://mac.r-project.org/tools/> \> Tools \> Mandatory tools \> *Install the tools that are pointed out in this page*

    -   For **Linux Users**: probably you know what you are doing where with high probability you will need to always **built from source locally**. If you need help I can assist you if you have installed Ubuntu and are stuck

# Data

The data collected by the corporation is available in the `sweep` package

-   Load the library: `library(sweep)`
-   Type `bike_sales` to access the data
-   If you want information about this data set type `?bike_sales`

The `bike_sales` data set contains 17 variables and a total of 15644 observations relate to transactional data:

```{r}
library(sweep)
library(tidyverse)
library(DT)

dim(bike_sales)

bike_sales |> 
  head() |> 
  datatable()
```

Each observation corresponds to a closed order. A closed order is a completed transaction where all items have been shipped and invoiced to a bicycle shop. A closed order, represented in the `order.id`, can be associated to 1 or more transactions and are capture by the `order.line` variable.

For example, the first 2 rows corresponds to a closed order where a bike shop named **Ithaca Mountain Climbers** bought 2 models: **Jekyll Carbon 2** and **Trigger Carbon 2**. The `order.id` is identify as `1` and the `order.line` variable has a value of `1` and `2` taking into account that `1` (see the `quantity` variable) **Jekyll Carbon 2** and `1` (see the `quantity` variable) **Trigger Carbon 2** model were acquired at a price of `6070` and `5970`.

```{r}
bike_sales |> 
  select(order.date, 
         order.id, order.line, 
         quantity, price,
         customer.id, bikeshop.name,
         model) |> 
  datatable()
```

You can use the following command inside RStudio IDE to inspect all the columns as a spreadsheet:

```{r}
#| eval: false
bike_sales |> 
  View()
```

# Marketing report structure part 3

In this third report you are going to use data from `bike_sales` as a third step to complete the 2 tasks by:

-   Estimating an statistical linear model to determine the price of a new product not offer by the company with information of the products the corporation already offer to support the Research and Development (R & D) division

-   Build a customer segmentation model by grouping the bike shops into clusters to support the marketing team

Use the template `003_marketing_report_part-3.qmd` to build the marketing report. You can find the template to use in 2 locations. Choose the option of your preference:

-   <https://github.com/luifrancgom/marketing_research_and_analytics/> \> 000_templates \> 003_marketing_report_part-3.qmd

-   Tercer Corte 40% \> Learning Resources \> Support material \> Templates \> 003_marketing_report_part-3.qmd

# References
