---
title: "Syllabus Marketing Research and Analytics Module A AE"
author:
  - name: Luis Francisco Gómez López
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Universidad Militar Nueva Granada
date: 2024-01-25
execute: 
  echo: false
  warning: false
format:
  html: 
    number-sections: true
    toc: true
    embed-resources: true
bibliography: ../000_references/marketing_research_and_analytics.bib
editor: visual
---

```{r libraries}
library(tidyverse)
library(kableExtra)
```

```{css}
p {
  text-align: justify;
}
```

# Professor and meeting hours with students

-   **Name**

    -   Luis Francisco Gomez López

-   **Communication through the virtual classroom**

    -   Every Saturday between 08:00 hours and 09:45 hours:

        -   *Communication channels \> Link Virtual meetings*

    -   For doubts and concerns of the topics of the course you have 2 options:

        -   *GitHub*:

            -   <https://github.com/luifrancgom/marketing_research_and_analytics> \> Issues \> New Issue

            -   You need to create a GitHub account if you don't have it: <https://github.com/> \> Sign up \> *Follow the instructions to create your personal account*

        -   *Forum: Questions or concerns* that can be accessed at the top of the platform

    -   Complaints regarding grades, please send an email to [**luis.gomezl\@unimilitar.edu.co**](mailto:luis.gomezl@unimilitar.edu.co){.email}

        -   To make any complaints, you are asked to wait first for the grades to be published on the News Bulletin Board and to be made based on the feedbacks that will be made taking into account the evaluation parameters

        -   Urgent issues, you can call me to +57 3124541536 Monday, Wednesday and Friday between 16:00 hours and 17:00 hours. Please do not send me whatsapp messages or call me on weekends.

# Presentation and objectives of the subject

## Description

-   Credits: 4
-   Type: Theoretical-Practice
-   Area of Knowledge: Professional
-   Semester: 1
-   Block: 1 (April 1 - June 4, 2024)

## Justification

The ability to analyze data and make decisions based on evidence is crucial in today's business world. Professionals in business administration need to be familiar with data analysis tools and techniques to better understand their customers, make informed decisions, and improve the profitability of their companies.

This course will provide students with an understanding of data analysis and marketing. The course will cover the basics of programming, statistics, and data analysis. Students will also learn how to apply these skills to solve problems in marketing and how to use a programming language to perform visualizations. By the end of the course, students will have the ability to use a programming language as a data analysis tool and apply their skills in a business context.

## General objective

Develop skills in programming, statistics, and data analysis to solve business problems related to marketing.

# Competencies

-   Global competencies:

    -   Apply statistical analysis techniques and programming in the evaluation of market data to identify trends and opportunities for improvement in products and commercial strategies.

-   Specific competencies:

    -   Prepare a reproducible market report by using statistical analysis techniques and programming to identify trends and opportunities for improvement in products and commercial strategies.

-   Competencies SABE curriculum

    -   Ability to use digital technologies effectively and efficiently
    -   Skills in handling a second language
    -   Ability to use the scientific method and logical thinking

# Schedule and contents

## Reading schedule

```{r}
# Include a footnote for the optional sections
# Generate a section for the exercises
# See https://github.com/quarto-dev/quarto-cli/issues/3340
tibble(Week = c(1:9),
       Material = rep("[@chapman_r_2019]", 
                      times = 9),
       Chapter = c(1:8, 11),
       Pages = c("1-12", 
                 "13-37, 41-43, 44-45",
                 "49-60, 62-69, 71-73, 75-77",
                 "79-101, 103-104, 106-109",
                 "111-132",
                 "133-145, 153-155",
                 "157-183, 185-189",
                 "193-206, 217-220",
                 "299-313, 336, 338-339"),
       Exercises = c("", 
                     "1-8",
                     "1-10",
                     "1-10, 12-15",
                     "1-7",
                     "1-9",
                     "1-12",
                     "1-7",
                     "1-5")) |> 
  kable()
```

### Primer corte 30%

-   April 1

    -   Start Block 2

    -   **Virtual meeting Presentation Syllabus**

        -   **Meeting time**: 20:00 - 20:45 hours
        -   **Prerequisites**:
            -   Review the course program in advance, which can be downloaded by following the route: **Primer corte 30% \> Learning Resources \> Support material \> Presentations \> 000_syllabus_marketing_rearch_and_analytics**
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 3

    -   **Welcome to R**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 1]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 5

    -   **Welcome to R**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 1]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 10

    -   **An overview of the R language**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 2]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 12

    -   **An overview of the R language**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 2]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 17

    -   **Describing data**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 3]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 19

    -   **Describing data**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 3]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 20

    -   **Oral presentation marketing report part 1**
        -   **Opening time**: 18:00 hours
        -   **Closing time**: 20:05 hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

### Segundo corte 30%

-   April 24

    -   **Relationship between continuous variables**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 4]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   April 26

    -   **Relationship between continuous variables**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 4]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 1

    -   **Comparing groups: tables and visualizations**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 5]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 3

    -   **Comparing groups: tables and visualizations**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 5]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 8

    -   **Comparing groups: statistical tests**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 6]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 10

    -   **Comparing groups: statistical tests**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 6]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 11

    -   **Oral presentation marketing report part 2**
        -   **Opening time**: 18:00 hours
        -   **Closing time**: 20:05 hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

### Tercer corte 40%

-   May 15

    -   **Identifying drivers of outcomes: linear models**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 7]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 17

    -   **Identifying drivers of outcomes: linear models**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 7]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 22

    -   **Reducing data complexity**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 8]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 24

    -   **Reducing data complexity**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 8]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 29

    -   **Segmentation: Clustering**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 11]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   May 31

    -   **Segmentation: Clustering**
        -   **Meeting time**: 19:00 - 20:30 hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 11]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   June 1

    -   **Oral presentation marketing report part 3**
        -   **Opening time**: 18:00 hours
        -   **Closing time**: 20:05 hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

# Qualifiable activities, weights and process in case of no presentation

## Activities and weights

-   **Primer corte**: 30%

    -   **Oral presentation marketing report part 1**: 30%

-   **Segundo corte**: 30%

    -   **Oral presentation marketing report part 2**: 30%

-   **Tercer corte**: 40%

    -   **Oral presentation marketing report part 3**: 40%

## Process in case of no presentation

-   Students who do not present an activity due to force majeure or illness, must send the respective excuse scanned through the [**luis.gomezl\@unimilitar.edu.co**](mailto:luis.gomezl@unimilitar.edu.co){.email}. If it is not sent, the note will be zero. Once the excuse has been satisfactorily received, the student should **not** present the activity and the corresponding grade will be equal to the immediately subsequent grade obtained in the next activity presented

# Software

## Install locally

-   R (programming language)

    -   <https://cloud.r-project.org/> \> Download and Install R \> *Choose your operating system and follow the instructions*

-   RStudio IDE (integrated development environment)

    -   Free desktop edition

        -   <https://posit.co/download/rstudio-desktop/> \> *Choose your operating system and follow the instructions*

-   Quarto (open-source scientific and technical publishing system)

    -   <https://quarto.org/docs/download/> \> Pre-release vX.X.X[^1] \> *Choose your operating system and follow the instructions*

[^1]: Quarto is constantly changing, so vX.X.X means the pre-release version that is available at the time of download

## Using a cloud-based solution

-   <https://posit.cloud/> \> Get Started \> Cloud Free \> Learn more \> Sign Up

    -   If you use this option after opening the account check out <https://posit.cloud/learn/guide> for getting started and using it

# Use of Chat GPT-3, GPT-4, Gemini 1.5, LLaMA 2 and its friends

You can use any artificial intelligence tool and in particular any Large Language Model where this practice is encourage when learning R or another programming language like Python or Julia. However, please use them in a responsible way by taking into account its capabilities and limitations and always pointing out to the professor and your colleges that you have use them.

If you have not used any Large Language Model you can try opening an account, for example in:

-   <https://chat.openai.com>
-   <https://gemini.google.com/>

Or you can run a Large Language Model locally, like CodeLlama 7B, using:

-   <https://lmstudio.ai/>

If you use the last option remember that you are using your personal computer so depending on the specifications of your machine you will or not will be able to use the chosen model

# Study material

## Principal material

-   [@chapman_r_2019, Chapters 1-8, 11]

    -   Complete route:

        -   <https://www.umng.edu.co/inicio> \> Biblioteca \> Bases virtuales y libros electrónicos \> S \> Springer Link \> Ingresar usuario y contraseña [^2] \> Search "R For Marketing Research and Analytics" \> R For Marketing Research and Analytics Book 2019 [^3]

    -   Direct link:

        -   <https://doi-org.ezproxy.umng.edu.co/10.1007/978-3-030-14316-9>

    -   Book companion website:

        -   <https://r-marketing.r-forge.r-project.org/>

            -   Code: <https://r-marketing.r-forge.r-project.org/code.html>
            -   Data: <https://r-marketing.r-forge.r-project.org/data.html>
            -   Exercices and solutions: <https://r-marketing.r-forge.r-project.org/exercises/index.html>

[^2]: Remember that the user and password to access and read the e-book are the same that you use to check your grades

[^3]: Download the book in your preferred format from the available options

## Support material

-   Quarto

    -   <https://quarto.org/>

        -   Get started with Rstudio: <https://quarto.org/docs/get-started/hello/rstudio.html>

        -   Guide: <https://quarto.org/docs/guide/> \> Documents \> HTML

        -   Reference: <https://quarto.org/docs/reference/> \> Documents \> HTML

-   StatQuest with Josh Starmer

    -   [https://www.youtube.com/\@statquest](https://www.youtube.com/@statquest){.uri}

        -   Video index

            -   <https://statquest.org/> \> Video Index

# About plagiarism

-   **Acuerdo 02 de 2015 (7 de Mayo): Artículo 68. Escala de Calificaciones Numeral f**

*"La comprobación de todo fraude o intento de fraude en cualquier tipo de prueba, acarreará al estudiante y a su(s) colaborador(es), una calificación de CERO PUNTO CERO (0.0) en la prueba respectiva. Como quiera que el fraude o intento de cometerlo es una falta grave, el docente debe informar de este hecho al Director del programa respectivo, quien enviará el informe a la Decanatura de la Facultad, con el fin de iniciar los trámites correspondientes para aplicar la sanción disciplinaria prevista en éste reglamento.”*

-   **Acuerdo 05 de 2018 (6 de Junio): Artículo 20 que modifica el Artículo 107 Faltas Disciplinarias Numeral c. Número 6 del Acuerdo 02 de 2015 (7 de Mayo)**

*"Hacer plagio en cualquier tipo de documento o desplegar alguna conducta dolosa que viole el régimen de propiedad intelectual de la Universidad."*

# References
