---
title: "Syllabus Marketing Research and Analytics Module A AE"
author:
  - name: Luis Francisco Gómez López
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Universidad Militar Nueva Granada
date: 2026-01-21
date-modified: last-modified
format:
  html: 
    number-sections: true
    toc: true
    embed-resources: true
execute: 
  echo: false
  warning: false
  message: false
params:
  meeting_info:
    value:
      phone: +57 3124541536
      days: [Monday, Wednesday, Friday]
      start_hour: 14:00
      end_hour: 15:00
  block:
    value:
      start_date: 2026-04-06
      end_date: 2026-06-07
  schedule:
    value:
      meeting_dates: [2, 4, 6]
      dates_recess: [2026-03-30, 2026-04-04]
      start_presentation_hour: 19:00
      end_presentation_hour: 21:00
      start_syllabus_hour: 20:00
      end_syllabus_hour: 20:45  
      start_virtual_meeting_hour: 08:00
      end_virtual_meeting_hour: 09:55
      start_activity_hour: 00:00
      end_activity_hour: 22:00
      start_exam_hour: 18:00
      end_exam_hour: 21:20
bibliography: ../000_references/marketing_research_and_analytics.bib
editor: source
---

```{r}
#| label: libraries

library(tidyverse)
library(glue)
library(knitr)
```

```{=html}
<style>
  p { 
    text-align: justify;
  }
</style>
```

# Instructor and meeting hours with students

```{r}
#| label: meeting-info

# Days ----
if (length(params$meeting_info$days) > 1) {
  days_str <- glue_collapse(params$meeting_info$days, sep = ", ", last = " or ")
} else {
  days_str <- params$meeting_info$days[1]
}
```

-   **Name**

    -   `r rmarkdown::metadata$author[[1]]$name`

-   **Communication through the virtual classroom**

    -   For doubts and concerns of the topics of the course, please use the *Forum: Questions or concerns* that can be accessed at the top of the platform

        -   In this way, doubts and concerns can be shared with all students and all course participants can contribute to the learning process

    -   Complaints regarding grades, please send an email to [**`{r} rmarkdown::metadata$author[[1]]$email`**](mailto:%60r%20rmarkdown::metadata$author%5B%5B1%5D%5D$email%60)

        -   Please wait for grades to be published on the Announcements forum before submitting any complaints. Complaints should be based on the feedback provided and must take into account the evaluation parameters.

    -   For urgent issues, you can call me to `{r} params$meeting_info$phone` between `{r} params$meeting_info$start_hour` hours and `{r} params$meeting_info$end_hour` hours on `{r} days_str`. But, please do not send me whatsapp messages or call me on weekends.

# Presentation and objectives of the subject

## Description

```{r}
#| label: block

# Dates ----
start_date <- ymd(params$block$start_date)
end_date <- ymd(params$block$end_date)

str_start_end_date <- paste(
  format(start_date, format = "%B %d"),
  " - ",
  format(end_date, format = "%B %d"),
  sep = "",
  ", ",
  year(start_date)
)
```

-   Credits: 4
-   Type: Theoretical-Practice
-   Area of Knowledge: Professional
-   Semester: 9
-   Block: 2 (`{r} str_start_end_date`)

## Justification

The ability to analyze data and make decisions based on evidence is crucial in today's business world. Professionals in business administration need to be familiar with data analysis tools and techniques to better understand their customers, make informed decisions, and improve the profitability of their companies.

This course will provide students with an understanding of data analysis and marketing. The course will cover the basics of programming, statistics, and data analysis. Students will also learn how to apply these skills to solve problems in marketing and how to use a programming language to perform visualizations. By the end of the course, students will have the ability to use a programming language as a data analysis tool and apply their skills in a business context.

## General objective

Develop skills in programming, statistics, and data analysis to solve business problems related to marketing.

## Learning outcomes

-   Course learning outcome:

    -   Apply statistical analysis techniques and programming in the evaluation of market data to identify trends and opportunities for improvement in products and commercial strategies.

-   Course specific learning outcomes:

    -   Prepare a reproducible market report by using statistical analysis techniques and programming to identify trends and opportunities for improvement in products and commercial strategies.

-   Competencies SABE curriculum

    -   Ability to use digital technologies effectively and efficiently
    -   Skills in handling a second language
    -   Ability to use the scientific method and logical thinking

# Schedule and contents

## Reading schedule

```{r}
#| label: tbl-reading-schedule
#| tbl-cap: Reading schedule

# Possible solution using kableExtra
## Include references
## See https://github.com/quarto-dev/quarto-cli/issues/3340
### #| results: asis
tibble(
  Week = c(1:9),
  Material = rep("[@chapman_r_2019]", times = 9),
  Chapter = c(1:8, 11),
  Pages = c(
    "1-12",
    "13-37, 41-43, 44-45",
    "49-60, 62-69, 71-73, 75-77",
    "79-101, 103-104, 106-109",
    "111-132",
    "133-145, 153-155",
    "157-183, 185-189",
    "193-206, 217-220",
    "299-313, 336, 338-339"
  ),
  Exercises = c(
    "",
    "1-8",
    "1-10",
    "1-10, 12-15",
    "1-7",
    "1-9",
    "1-12",
    "1-7",
    "1-5"
  )
) |>
  kable()
# kbl() |>
# kable_styling() |>
# unclass() |>
# cat()
```

## Detailed schedule

```{r}
#| label: detailed-schedule

source(file = "../000_scripts/000_tbl_schedule.R")

tbl_schedule <- tbl_schedule(
  start_date = params$block$start_date,
  end_date = params$block$end_date,
  dates_recess = params$schedule$dates_recess,
  meeting_days = params$schedule$meeting_dates
)

virtual_meeting <- tbl_schedule |>
  filter(virtual_meeting == 1) |>
  pull(index)

activity <- tbl_schedule |>
  filter(activity == 1) |>
  pull(index)

exam <- tbl_schedule |>
  filter(exam == 1) |>
  pull(index)
```

### Primer corte 30%

-   `r format(x = ymd(params$block$start_date), format = "%B %e")`

    -   **Virtual meeting Presentation Syllabus**
        -   **Meeting time**: `r paste(params$schedule$start_syllabus_hour, params$schedule$end_syllabus_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Review the course program in advance, which can be downloaded by following the route *Primer corte 30% \> Learning Resources \> Support material \> Presentations \> 000_syllabus_marketing_research_and_analytics\_...*
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[1]], format = "%B %e")`

    -   **Welcome to R**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 1]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[2]], format = "%B %e")`

    -   **Business Case**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[3]], format = "%B %e")`

    -   **Progress marketing report 1**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[4]], format = "%B %e")`

    -   **An overview of the R language**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 2]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[5]], format = "%B %e")`

    -   **An overview of the R language**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 2]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[6]], format = "%B %e")`

    -   **Progress marketing report 1**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[7]], format = "%B %e")`

    -   **Describing data**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 3]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[8]], format = "%B %e")`

    -   **Describing data**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 3]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[9]], format = "%B %e")`

    -   **Progress marketing report 1**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = exam[[1]], format = "%B %e")`

    -   **Oral presentation marketing report part 1**
        -   **Opening time**: `r paste(params$schedule$start_exam_hour)` hours
        -   **Closing time**: `r paste(params$schedule$end_exam_hour)` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

### Segundo corte 30%

<!-- -   `r format(x = ymd(params$schedule$dates_recess[[1]]), format = "%B %e")`

    -   **Begin Holy Week**

-   `r format(x = ymd(params$schedule$dates_recess[[2]]), format = "%B %e")`

    -   **End Holy Week** -->

-   `r format(x = virtual_meeting[[10]], format = "%B %e")`

    -   **Relationship between continuous variables**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 4]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[11]], format = "%B %e")`

    -   **Relationship between continuous variables**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 4]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[12]], format = "%B %e")`

    -   **Progress marketing report 2**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[13]], format = "%B %e")`

    -   **Comparing groups: tables and visualizations**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 5]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[14]], format = "%B %e")`

    -   **Comparing groups: tables and visualizations**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 5]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[15]], format = "%B %e")`

    -   **Progress marketing report 2**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[16]], format = "%B %e")`

    -   **Comparing groups: statistical tests**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 6]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[17]], format = "%B %e")`

    -   **Comparing groups: statistical tests**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 6]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[18]], format = "%B %e")`

    -   **Progress marketing report 2**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = exam[[2]], format = "%B %e")`

    -   **Oral presentation marketing report part 2**
        -   **Opening time**: `r paste(params$schedule$start_exam_hour)` hours
        -   **Closing time**: `r paste(params$schedule$end_exam_hour)` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

### Tercer corte 40%

-   `r format(x = virtual_meeting[[19]], format = "%B %e")`

    -   **Identifying drivers of outcomes: linear models**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 7]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[20]], format = "%B %e")`

    -   **Identifying drivers of outcomes: linear models**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 7]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[21]], format = "%B %e")`

    -   **Progress marketing report 3**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[22]], format = "%B %e")`

    -   **Reducing data complexity**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 8]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[23]], format = "%B %e")`

    -   **Reducing data complexity**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 8]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[24]], format = "%B %e")`

    -   **Progress marketing report 3**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = virtual_meeting[[25]], format = "%B %e")`

    -   **Segmentation: Clustering**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 9]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[26]], format = "%B %e")`

    -   **Segmentation: Clustering**
        -   **Meeting time**: `r paste(params$schedule$start_presentation_hour, params$schedule$end_presentation_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Read and review [@chapman_r_2019, Chapter 9]
            -   Have a headset with a built-in microphone to isolate noise from the environment

-   `r format(x = virtual_meeting[[27]], format = "%B %e")`

    -   **Progress marketing report 3**
        -   **Meeting time**: `r paste(params$schedule$start_virtual_meeting_hour, params$schedule$end_virtual_meeting_hour, sep = " - ")` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment
        -   **Observations**:
            -   At least 1 member of the group must attend
            -   The meeting will last 10 minutes per group, no more and no less

-   `r format(x = exam[[3]], format = "%B %e")`

    -   **Oral presentation marketing report part 3**
        -   **Opening time**: `r paste(params$schedule$start_exam_hour)` hours
        -   **Closing time**: `r paste(params$schedule$end_exam_hour)` hours
        -   **Prerequisites**:
            -   Have a headset with a built-in microphone to isolate noise from the environment

# Qualifiable activities, weights and process in case of no presentation

## Activities and weights

-   **Primer corte (3 weeks)**: 30%

    -   **Oral presentation marketing report part 1**: 30%

-   **Segundo corte (3 weeks)**: 30%

    -   **Oral presentation marketing report part 2**: 30%

-   **Tercer corte (3 weeks)**: 40%

    -   **Oral presentation marketing report part 3**: 40%

[Disclaimer:]{style="color: red;"} Participants enrolled in the course will work in groups in a business case taken from the course [Data Science for Business Part 1](https://university.business-science.io/p/ds4b-101-r-business-analysis-r) offered by the company [Business Science](https://www.business-science.io/) and adapted to be in line with the topics covered in [@chapman_r_2019].

## Process in case of no presentation

-   Students who are unable to attend an oral presentation due to force majeure or illness must submit a scanned copy of their official excuse to [**luis.gomezl\@unimilitar.edu.co**](mailto:luis.gomezl@unimilitar.edu.co){.email}. Failure to provide a valid excuse will result in a zero grade for the presentation.

    -   If the excuse is approved, the absent student will be exempt from attending the oral presentation. Their grade will depend on their contribution to the group project:

        -   If the group confirms the absent student actively participated in the project's development throughout the designated three-week period, they will receive the same grade as the rest of the group.

        -   If the student was unable to participate due to the force majeure or illness, their grade will be equal to the grade they receive on the subsequent activity. If the absence occurred during the final oral presentation, their grade will be equal to the grade they receive on the second oral presentation. To receive this benefit, the student must actively participate in the development of at least 2 oral presentations.

# Software

-   Check out all the necessary instructions in:

    -   [Primeros pasos curso Marketing Research and Analytics con R](https://youtube.com/playlist?list=PLvL0Mlbw5cm0-BI4LfX76OpgcNwLQNZSW&si=YzBoowOavXzmqMxq)

## Install locally

-   R (programming language)

    -   <https://cloud.r-project.org/> \> Download and Install R \> *Choose your operating system and follow the instructions*

-   RStudio IDE (integrated development environment)

    -   Free desktop edition

        -   <https://posit.co/download/rstudio-desktop/> \> *Choose your operating system and follow the instructions*

-   Quarto (open-source scientific and technical publishing system)

    -   <https://quarto.org/docs/download/> \> Current Release vX.X.X[^1] \> *Choose your operating system and follow the instructions*

[^1]: Quarto is constantly changing, so vX.X.X means the current Release version that is available at the time of download

## Using a cloud-based solution

-   <https://posit.cloud/> \> Get Started \> Cloud Free \> Learn more \> Sign Up

    -   If you use this option after opening the account check out <https://posit.cloud/learn/guide> for getting started and using it

# Use of Chat GPT-4, Gemini 1.5, Copilot and its friends

In this course, the use of generative artificial intelligence tools is allowed as long as you follow the following good practices [@equipo_proyecto_ia-uniandes_lineamientos_2024, pp. 20-21]:

-   **Informed use**: learn beforehand how the tool works, its limitations, and risks
-   **Transparent use**: report which tool you used and how it was used it
-   **Ethical use**: do not present material generated by the tool as your own and follow the corresponding citation rules (see [@mcadoo_how_2024]).
-   **Critical evaluation**: always contrast and compare any response obtained through the tool with reliable or primary sources of information
-   **Responsible use**: do not include personal or confidential information (own or others) when entering instructions (prompts)

For more detailed guidance on good practices and recommended uses, check out [@equipo_proyecto_ia-uniandes_lineamientos_2024, Section 3.2]

# Study material

## Principal material

-   [@chapman_r_2019, Chapters 1-8, 11]

    -   <a href="https://doi-org.ezproxy.umng.edu.co/10.1007/978-3-030-14316-9" target="_blank" rel="noopener noreferrer"> https://doi-org.ezproxy.umng.edu.co/10.1007/978-3-030-14316-9</a>

    -   Book companion website:

        -   <https://r-marketing.r-forge.r-project.org/>

## Support material

-   Quarto

    -   <https://quarto.org/>

        -   Get started with Rstudio: <https://quarto.org/docs/get-started/hello/rstudio.html>

        -   Guide: <https://quarto.org/docs/guide/> \> Documents \> HTML

        -   Reference: <https://quarto.org/docs/reference/> \> Documents \> HTML

-   StatQuest with Josh Starmer

    -   [https://www.youtube.com/\@statquest](https://www.youtube.com/@statquest){.uri}

        -   Video index

            -   <https://statquest.org/> \> Video Index

# About plagiarism

-   **Acuerdo 02 de 2015 (7 de Mayo): Artículo 68. Escala de Calificaciones Numeral f**

*"La comprobación de todo fraude o intento de fraude en cualquier tipo de prueba, acarreará al estudiante y a su(s) colaborador(es), una calificación de CERO PUNTO CERO (0.0) en la prueba respectiva. Como quiera que el fraude o intento de cometerlo es una falta grave, el docente debe informar de este hecho al Director del programa respectivo, quien enviará el informe a la Decanatura de la Facultad, con el fin de iniciar los trámites correspondientes para aplicar la sanción disciplinaria prevista en éste reglamento.”*

-   **Acuerdo 05 de 2018 (6 de Junio): Artículo 20 que modifica el Artículo 107 Faltas Disciplinarias Numeral c. Número 6 del Acuerdo 02 de 2015 (7 de Mayo)**

*"Hacer plagio en cualquier tipo de documento o desplegar alguna conducta dolosa que viole el régimen de propiedad intelectual de la Universidad."*

# References
