---
title: "Marketing report part 3"
date: 2025-11-29
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gómez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| message: false

library(sweep)
library(tidyverse)
library(scales)
library(tidymodels)
library(DT)
```

# Modeling

## Introduction

Pending

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales |> 
  mutate(order.id = factor(x = order.id, 
                           ordered = TRUE),
         order.line = factor(x = order.line,
                             ordered = TRUE),
         customer.id = factor(x = customer.id,
                              ordered = FALSE),
         bikeshop.name = factor(x = bikeshop.name,
                                ordered = FALSE),
         bikeshop.city = factor(x = bikeshop.city,
                                ordered = FALSE),
         bikeshop.state = factor(x = bikeshop.state,
                                 ordered = FALSE),
         product.id = factor(x = product.id,
                             ordered = FALSE),
         model = factor(x = model,
                        ordered = FALSE),
         category.primary = factor(category.primary,
                                   ordered = FALSE),
         category.secondary = factor(x = category.secondary,
                                     ordered = FALSE),
         frame = factor(x = frame,
                        ordered = FALSE))
```

## The product gap

Pending

```{r}
#| message: false

# Grouping ----
total_revenue_by_model_cat2_frame <- bike_sales |> 
  group_by(model, 
           category.secondary,
           frame) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  ungroup() |> 
  mutate(category.secondary = fct_reorder(
      .f = category.secondary,
      .x = total_revenue
  ))
```

```{r}
# Visualization ----
total_revenue_by_model_cat2_frame |> 
  ggplot() + 
  geom_point(
    aes(x = frame, y = total_revenue),
    position = position_jitter(height = 0.2,
                               width = 0.2)
  ) +
  scale_y_continuous(
    labels = label_currency()
  ) +
  facet_wrap(
    facets = vars(category.secondary)
  ) +
  labs(
    x = "Frame",
    y = "Total revenue (US dollars)"
  )
```

## Determine the price of a new product

Pending

### Selecting a new potential product

Pending: choose the model to produce

Bla bla bla

-   Cyclocross Aluminum
-   Fat Bike Carbon
-   Over Mountain Aluminum
-   Sport Carbon
-   Triathalon Aluminum

### Building a model to price a new product

Pending

#### Choosing a model

Pending

-   `price ~ category.secondary + frame + bikeshop.state`

#### Fit the model

Pending

```{r}
model_4 <- lm(
  formula = price ~ category.secondary + frame + bikeshop.state,
  data = bike_sales
)

model_4_tidy <- model_4 |> 
  tidy()

model_4_tidy |> 
  datatable(
    colnames = c("Term",
                 "Estimate",
                 "Standard Error",
                 "Statistic",
                 "P-value") 
  ) |> 
  formatRound(
    columns = c("estimate",
                "std.error",
                "statistic",
                "p.value"), 
    digits = 3  
      )
```

#### Predict the price of the new product

Pending

```{r}
new_data <- tibble(
  category.secondary = c("Fat Bike",
                         "Fat Bike"),
  frame = c("Carbon",
            "Carbon"),
  bikeshop.state = c("CA",
                     "PA")
)

model_4_pred <- predict(
  object = model_4,
  newdata = new_data
)

model_4_pred |> 
  enframe(
    name = "obsevation",
    value = "pred_price"
  ) |> 
  bind_cols(
    new_data
  ) |> 
  datatable(
    colnames = c("Observation",
                 "Predict Price",
                 "Category Secondary",
                 "Frame",
                 "State")
    ) |> 
  formatRound(
    columns = c("pred_price"),
    digits = 2
  )
```

## Segment bike shops with k-means clustering

Add some text pointing out the principal findings as a result of applying a k-means clustering model to segment the bike shops

### Bike shop revenue trends

In this section the objective is to group and aggregate data to calculate the total revenue using the variable `price.ext`

-   Select the variables `bikeshop.name`, `price.ext`, `model`, `category.primary`, `category.secondary` and `frame`

-   Group the data by `bikeshop.name`, `model`, `category.primary`, `category.secondary` and `frame`

-   Calculate the total revenue and ungroup the data using the `ungroup` function or the `.groups = "drop"` argument from `summarise`

```{r}
# In this part you can use the functions
# select, group_by and summarise
## You can use at the end ungroup of 
## the argument .groups = "drop" to 
## ungroup the data
### If you name the variable as total_revenue the first 5 rows 
### and the first and last column of the  final data will be:

# # A tibble: 5 × 2
#   bikeshop.name      total_revenue
#   <chr>                      <dbl>
# 1 Albuquerque Cycles         92800
# 2 Albuquerque Cycles          5320
# 3 Albuquerque Cycles          8310
# 4 Albuquerque Cycles         19170
# 5 Albuquerque Cycles          1620

### Remember that you will get 6 columns but I am only showing
### 2 columns of the data
```

### Prepare data

The objective of this part is to:

-   Normalize the data so you can apply latter the **principal component analysis** dimensionality reduction technique

-   Create a customer-product table with a column that includes all the bikeshops and the rest of the columns with information about the percentage of total revenue of the **97** models offered by the corporation and described in the variable `model`

#### Normalization

```{r}
# Follow this template to normalize the data

# bike_sales_total_revenue_pct <- bike_sales_total_revenue |> 
#   group_by(bikeshop.name) |> 
#   mutate(total_revenue_pct = total_revenue / sum(total_revenue)) |> 
#   ungroup()

## You can check the data using
### bike_sales_total_revenue_pct |> glimpse()
```

#### Customer-product table

In this part you are going to use the function `pivot_wider` to create the customer-product table. For more information checkout `?pivot_wider`

Also check out the argument `values_fill` from this function and try this function with and without this argument so you can understand why it is important to use `values_fill = 0`

```{r}
# Follow this template to create a customer-product table

# customer_product_table <- bike_sales_total_revenue_pct |> 
#   select(bikeshop.name,
#          model,
#          total_revenue_pct) |> 
#   pivot_wider(id_cols = bikeshop.name, 
#               names_from = model, 
#               values_from = total_revenue_pct, 
#               values_fill = 0)

## You can check the data using
### bike_sales_total_revenue_pct |> glimpse()

#### A particular cell of the data will correspond to 
#### the percentage revenue generated by a model and a 
#### bike shop
#### For example if the row and the column corresponds to
##### bikeshop.name = Albuquerque Cycles
##### `Bad Habit 1` = 0.00571
#### It will mean that this model has a participation 
#### in the total revenue generated by Albuquerque 
#### Cycles of 0.571 %
```

### Performing k-means clustering

The objective of this section is to apply the k-means clustering technique by grouping the bike shops

-   First using the function `n_clusters` from the package `parameters` on the customer-product table to identify the number of clusters to use for k-means

```{r}
# Follow this template to identify the number of clusters to use in k-means

# library(parameters)

# n_clusters_customer_product_table <- customer_product_table |> 
#   select(-bikeshop.name) |> 
#   n_clusters(standardize = FALSE,
#              nbclust_method = "kmeans")

# n_clusters_customer_product_table

# If you want, you can check the 
# criteria or methods used to select 
# the number of clusters to use in kmeans
# This part is optional

# n_clusters_customer_product_table |> 
#   as_tibble()
```


Use the `kmeans` function with the argument `centers` equal to the number of clusters identify above and applying it to the customer product table

-   Remember to use `set.seed` to fix the initial centers. For example, you can use `set.seed(seed = 1234)`  

-   Remember that `kmeans` only works with numerical data so you will need to drop the `bikeshop.name` from the customer product table

    -   Checkout `?kmeans` and the argument `x` in relation to this aspect

-   Once you have apply the k-means clustering technique save this object as `kmeans_object`

```{r}
# Follow this template to apply k-means
# to the customer-product table

# set.seed(seed = 1234)
# kmeans_object <- customer_product_table |> 
#   select(-bikeshop.name) |> 
#   kmeans(centers = 4)
```

#### Extracting the cluster assignment

Now that you have apply the k-means clustering technique you are going to extract the clusters and assign them to each bike shop

You are going to use the `glance` function from the `tidymodels` library to extract the cluster assignment and select the bike shop names and the clusters to obtain the information required by the marketing team

```{r}
# Follow this template to obtain the relevant information
# required by the marketing team

# clusters <- kmeans_object |> 
#   augment(data = customer_product_table) |> 
#   select(bikeshop.name, .cluster)

## You can check the data using
### clusters |> glimpse()
### clusters |> count(.cluster)
```

### Data visualization with k-means clustering and principal component analysis

It is difficult to visualize the clusters that are identified using `kmeans` taking into account that in our case we use **97** variables (the participation of the total revenue of the **97** models) and a human being can not visualize a plot 97 dimensional plot

In this section you are going to reduce the data complexity by generating a 2 dimensional projection of the **97** variables by applying the principal component analysis technique

Use the `prcomp` function with the argument `center = TRUE` and applying it to the customer product table

-   Remember that `prcomp` only works with numerical data so you will need to drop the `bikeshop.name` from the customer product table

    -   Checkout `?prcomp` and the argument `x` in relation to this aspect

-   Also take into account that you don't need to scale the data using the argument `scale. = TRUE` because each data point from the customer product table corresponds to the participation of the total revenue of each model generated by each bike shop

-   Once you have apply the principal component analysis technique save this object as `pca_object`

```{r}
# Follow this template to apply the principal component analysis technique

# pca_object <- customer_product_table |>
#   select(-bikeshop.name) |> 
#   prcomp(center = TRUE, 
#          scale. = FALSE)
```

#### Extracting the first 2 principal components

In this section you are going to extract the first 2 principal components so you can use them to visualize the clusters in a 2 dimensional plot

```{r}
# Follow this template to obtain the first 2 
# principal components

# pca_1_2 <- pca_object$x |> 
#   as_tibble() |> 
#   select(PC1, PC2)

## Name this object as pca_1_2 so you will remember
## that it contains the first 2 principal components

## You can check the data using
### pca_1_2 |> glimpse()
```

#### Cluster visualization using k-means and principal component analysis

Using the information about the cluster assignment and the 2 principal components consolidate these parts into a tibble to generate a plot and visualize the clusters using the 2 dimensional projection

```{r}
# Follow this template to visualize the fill clusters

# clusters_pca_1_2 <- pca_1_2 |> 
#   bind_cols(clusters)

# clusters_pca_1_2 |> 
#   ggplot() + 
#   geom_point(aes(x = PC1, 
#                  y = PC2,
#                  color = .cluster))

## Improve the plot by adding labels to the x, y axis 
## and a title to explain the visualization to the
## marketing team
```
