---
title: "Marketing report part 3"
date: 2025-05-31
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gomez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: visual
---

```{r }
#| label: libraries
#| echo: false
#| message: false

library(tidyverse)
library(sweep)
library(DT)
library(tidymodels)
```

# Modeling

## Introduction

Pending

## Import data

Pending

```{r}
# Import data ----
bike_sales <- bike_sales

# Transform data ----
bike_sales <- bike_sales |> 
  mutate(order.id = factor(x = order.id, 
                           ordered = FALSE),
         order.line = factor(x = order.line, 
                             ordered = TRUE),
         customer.id = factor(x = customer.id, 
                              ordered = FALSE),
         bikeshop.name = factor(x = bikeshop.name, 
                                ordered = FALSE),
         bikeshop.city = factor(x = bikeshop.city, 
                                ordered = FALSE),
         bikeshop.state = factor(x = bikeshop.state, 
                                 ordered = FALSE),
         product.id = factor(x = product.id, 
                             ordered = FALSE),
         model = factor(x = model, 
                        ordered = FALSE),
         category.primary = factor(x = category.primary, 
                                   ordered = FALSE),
         category.secondary = factor(x = category.secondary, 
                                     ordered = FALSE),
         frame = factor(x = frame, 
                        ordered = FALSE))
```

## The product gap

Pending

@tbl-total-revenue-by-cat2-frame

```{r}
#| label: tbl-total-revenue-by-cat2-frame
#| tbl-cap: Total revenue by category secondary and frame
#| message: false

total_revenue_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(revenue = sum(price.ext)) |> 
  ungroup() |> 
  arrange(desc(revenue)) |> 
  mutate(category.secondary = fct_reorder(.f = category.secondary,
                                          .x = revenue))

total_revenue_by_cat2_frame |> 
  datatable(colnames = c("Category secondary",
                         "Frame",
                         "Revenue")) 
```

Pending

@fig-total-revenue-by-cat2-frame

```{r}
#| label: fig-total-revenue-by-cat2-frame
#| fig-cap: Total revenue by category secondary and frame

## Visualization ----
total_revenue_by_cat2_frame |> 
  ggplot() + 
  geom_point(aes(x = frame,
                 y = revenue),
             position = position_jitter(width = 0.2,
                                        height = 0.2)) +
  scale_y_continuous(labels = scales::label_currency()) + 
  facet_wrap(facets = vars(category.secondary)) +
  labs(x = "Frame",
       y = "Revenue (US Dollars)")
```

## Determine the price of a new product

Pending

### Selecting a new potential product

Pending

-   Cyclocross Aluminum
-   Fat Bike Carbon
-   Over Mountain Aluminum
-   Sport Carbon
-   Triathalon Aluminum

Pending

### Building a model to price a new product

Pending

#### Choosing a model

Pending

-   `price ~ category.secondary + frame`

**Note 1**: If the model is to complex like `price ~ category.secondary*frame*bikeshop.state` you will obtain a considerable parameters but if the model is to simple like `price ~ category.secondary` you will not be able to obtain good predictions. Choose the level of complexity that you consider appropriate to help the Research and Development (R & D) division and to determine the new price of the new product

**Note 2**: If you find that the model generates silly results, like for example negative or 0 values for the price, abandon the choosen model and use another model or change the product gap combination or discard particular locations where the product will going to be offer.

#### Fit the model

Pending

@tbl-model-2

```{r}
#| label: tbl-model-2
#| tbl-cap: Linear model for price prediction of new products

## Model 2 ----
model2 <- lm(formula = price ~ category.secondary + frame,
             data = bike_sales)

model_tidy2 <- model2 |> 
  tidy()

model_tidy2 |> 
  datatable(colnames = c("Term",
                         "Estimation",
                         "Standard Error",
                         "T-statistic",
                         "P-value")) |> 
  formatRound(c("estimate", 
                "std.error", 
                "statistic", 
                "p.value"))
```

#### Predict the price of the new product

Pending

```{r}
#| label: tbl-model2-price-pred
#| tbl-cap: Price prediction of new products

## Predictions ----
new_data <- tibble(category.secondary = c("Cyclocross",
                                          "Fat Bike",
                                          "Over Mountain",
                                          "Sport",
                                          "Triathalon"),
                   frame = c("Aluminum",
                             "Carbon",
                             "Aluminum",
                             "Carbon",
                             "Aluminum"))

predict_prices_model2 <- predict(object = model2, 
        newdata = new_data) |> 
  enframe(name = "observations",
          value = "pred_price") |> 
  bind_cols(new_data) |> 
  arrange(desc(pred_price))

predict_prices_model2 |> 
  select(-c(observations)) |> 
  datatable(colnames = c("Price prediction",
                         "Category secondary",
                         "Frame")) |> 
  formatRound(c("pred_price"))
```

## Segment bike shops with k-means clustering

Add some text pointing out the principal findings as a result of applying a k-means clustering model to segment the bike shops

### Bike shop revenue trends

In this section the objective is to group and aggregate data to calculate the total revenue using the variable `price.ext`

-   Select the variables `bikeshop.name`, `price.ext`, `model`, `category.primary`, `category.secondary` and `frame`

-   Group the data by `bikeshop.name`, `model`, `category.primary`, `category.secondary` and `frame`

-   Calculate the total revenue and ungroup the data using the `ungroup` function or the `.groups = "drop"` argument from `summarise`

```{r}
# In this part you can use the functions
# select, group_by and summarise
## You can use at the end ungroup of 
## the argument .groups = "drop" to 
## ungroup the data
### If you name the variable as total_revenue the first 5 rows 
### and the first and last column of the  final data will be:

# # A tibble: 5 × 2
#   bikeshop.name      total_revenue
#   <chr>                      <dbl>
# 1 Albuquerque Cycles         92800
# 2 Albuquerque Cycles          5320
# 3 Albuquerque Cycles          8310
# 4 Albuquerque Cycles         19170
# 5 Albuquerque Cycles          1620

### Remember that you will get 6 columns but I am only showing
### 2 columns of the data
```

### Prepare data

The objective of this part is to:

-   Normalize the data so you can apply latter the **principal component analysis** dimensionality reduction technique

-   Create a customer-product table with a column that includes all the bikeshops and the rest of the columns with information about the percentage of total revenue of the **97** models offered by the corporation and described in the variable `model`

#### Normalization

```{r}
# Follow this template to normalize the data

# bike_sales_total_revenue_pct <- bike_sales_total_revenue |> 
#   group_by(bikeshop.name) |> 
#   mutate(total_revenue_pct = total_revenue / sum(total_revenue)) |> 
#   ungroup()

## You can check the data using
### bike_sales_total_revenue_pct |> glimpse()
```

#### Customer-product table

In this part you are going to use the function `pivot_wider` to create the customer-product table. For more information checkout `?pivot_wider`

Also check out the argument `values_fill` from this function and try this function with and without this argument so you can understand why it is important to use `values_fill = 0`

```{r}
# Follow this template to create a customer-product table

# customer_product_table <- bike_sales_total_revenue_pct |> 
#   select(bikeshop.name,
#          model,
#          total_revenue_pct) |> 
#   pivot_wider(id_cols = bikeshop.name, 
#               names_from = model, 
#               values_from = total_revenue_pct, 
#               values_fill = 0)

## You can check the data using
### bike_sales_total_revenue_pct |> glimpse()

#### A particular cell of the data will correspond to 
#### the percentage revenue generated by a model and a 
#### bike shop
#### For example if the row and the column corresponds to
##### bikeshop.name = Albuquerque Cycles
##### `Bad Habit 1` = 0.00571
#### It will mean that this model has a participation 
#### in the total revenue generated by Albuquerque 
#### Cycles of 0.571 %
```

### Performing k-means clustering

The objective of this section is to apply the k-means clustering technique by grouping the bike shops in **5** clusters

-   If you want you can try with a different number of clusters but first it is important that you at least be able to complete the task with **5** clusters

Use the `kmeans` function with the argument `centers = 5` and applying it to the customer product table

-   Remember that `kmeans` only works with numerical data so you will need to drop the `bikeshop.name` from the customer product table

    -   Checkout `?kmeans` and the argument `x` in relation to this aspect

-   Once you have apply the k-means clustering technique save this object as `kmeans_object`

#### Extracting the cluster assignment

Now that you have apply the k-means clustering technique you are going to extract the clusters and assign them to each bike shop

You are going to use the `glance` function from the `tidymodels` library to extract the cluster assignment and select the bike shop names and the clusters to obtain the information required by the marketing team

```{r}
# Follow this template to obtain the relevant information
# required by the marketing team

# clusters <- kmeans_object |> 
#   augment(data = customer_product_table) |> 
#   select(bikeshop.name, .cluster)

## You can check the data using
### clusters |> glimpse()
### clusters |> count(.cluster)
```

### Data visualization with k-means clustering and principal component analysis

It is difficult to visualize the clusters that are identified using `kmeans` taking into account that in our case we use **97** variables (the participation of the total revenue of the **97** models) and a human being can not visualize a plot 97 dimensional plot

In this section you are going to reduce the data complexity by generating a 2 dimensional projection of the **97** variables by applying the principal component analysis technique

Use the `prcomp` function with the argument `center = TRUE` and applying it to the customer product table

-   Remember that `prcomp` only works with numerical data so you will need to drop the `bikeshop.name` from the customer product table

    -   Checkout `?prcomp` and the argument `x` in relation to this aspect

-   Also take into account that you don't need to scale the data using the argument `scale. = TRUE` because each data point from the customer product table corresponds to the participation of the total revenue of each model generated by each bike shop

-   Once you have apply the principal component analysis technique save this object as `pca_object`

#### Extracting the first 2 principal components

In this section you are going to extract the first 2 principal components so you can use them to visualize the clusters in a 2 dimensional plot

```{r}
# Follow this template to obtain the first 2 
# principal components

# pca_1_2 <- pca_object$x |> 
#   as_tibble() |> 
#   select(PC1, PC2)

## Name this object as pca_1_2 so you will remember
## that it contains the first 2 principal components

## You can check the data using
### pca_1_2 |> glimpse()
```

#### Cluster visualization using k-means and principal component analysis

Using the information about the cluster assignment and the 2 principal components consolidate these parts into a tibble to generate a plot and visualize the clusters using the 2 dimensional projection

```{r}
# Follow this template to visualize the fill clusters

# clusters_pca_1_2 |> 
#   ggplot() + 
#   geom_point(aes(x = PC1, 
#                  y = PC2,
#                  color = .cluster))

## Improve the plot by adding labels to the x, y axis 
## and a title to explain the visualization to the
## marketing team
```
