---
title: "Describing data"
author: 
  - Norah Jones
  - Bill Gates
  - Luis Francisco Gómez
format: 
  html:
    embed-resources: true
    toc: true
    number-sections: true
    code-fold: true
editor: visual
---

```{r}
#| label: libraries
#| message: false

library(tidyverse)
library(skimr)
library(DT)
```

# Introduction

In @fig-plot-weekly-store-by-country you can see the Germany is the country with the greatest sales of product 2.

# Import

The data used is from a set of stores in different parts of the world.

```{r}
#| message: false

weekly_store <- read_csv(file = "../000_data/003_data_chapter3.csv")
```

# Transform data

```{r}
weekly_store <- weekly_store |> 
  mutate(storeNum = factor(x = storeNum, ordered = FALSE),
         Year = factor(x = Year, levels = 1:2, ordered = TRUE),
         Week = factor(x = Week, levels = 1:52, ordered = TRUE),
         p1prom = as.logical(x = p1prom),
         p2prom = as.logical(x = p2prom))
```

# Summarize data

```{r}
weekly_store |> 
  skim()
```

# Answering questions with data

## Prepare data

In @tbl-weekly-store-by-country you can see the different sales by country of product 2 

```{r}
#| label: tbl-weekly-store-by-country
#| tbl-cap: "Total sales by country of product 2"

weekly_store_by_country <- weekly_store |> 
  group_by(country) |> 
  summarise(sum_p2sales = sum(p2sales)) |> 
  mutate(country = fct_reorder(.f = country, 
                               .x = sum_p2sales))

weekly_store_by_country |> 
  datatable()
```

## Data visualization

```{r}
#| label: fig-plot-weekly-store-by-country
#| fig-cap: "Total sales by country of product 2"

weekly_store_by_country |> 
  ggplot() +
  geom_col(aes(x = country, 
               y = sum_p2sales,
               fill = country),
           color = "black") + 
  labs(x = "Country",
       y = "Total sales",
       title = "Total sales of product 2 by country")
```
