---
title: "Marketing report part 3"
date: 2026-06-06
author: 
  - Norah Jones   
  - Bill Gates
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| echo: false
#| warning: false

library(sweep)
library(tidyverse)
library(scales)
library(tidymodels)
library(DT)
library(tidyheatmaps)
library(parameters)
```

# Modeling

## Introduction

Pending

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## The product gap

Pending

### Prepare data

Pending

```{r}
#| message: false
total_sales_by__model_cat2_frame <- bike_sales |> 
  select(
    model,
    category.secondary,
    frame,
    price.ext
  ) |> 
  group_by(
    model, category.secondary, frame
  ) |> 
  summarise(
    total_revenue = sum(price.ext)
  ) |> 
  ungroup() |> 
  arrange(desc(total_revenue)) |> 
  mutate(category.secondary = fct_reorder(
    .f = category.secondary,
    .x = total_revenue
  )
)
```

### Visualization

Pending

```{r}
#| label: fig-total-sales-by-model-cat2-frame
#| fig-cap: Product gap and total sales by model, category secondary and frame
total_sales_by__model_cat2_frame |> 
  ggplot() + 
  geom_point(
    aes(x = frame, y = total_revenue),
    position = position_jitter(
      width = 0.2,
      height = 0.2
    )
  ) +
  scale_y_continuous(
    labels = label_currency()
  ) +
  facet_wrap(
    facets = vars(category.secondary)
  ) +
  labs(
    x = "Frame",
    y = "Total revenue (US dollars)"
  )
```

## Determine the price of a new product

Pending

### Selecting a new potential product

Pending

-   Cyclocross Aluminum
-   Fat Bike Carbon
-   Over Mountain Aluminum
-   Sport Carbon
-   Triathalon Aluminum

**Note**: choose 1 product from the product gap

### Building a model to price a new product

Pending

#### Choosing a model

Pending

-   `price ~ category.secondary + frame`

**Note 1**: If the model is to complex like `price ~ category.secondary*frame*bikeshop.state` you will obtain a considerable number of parameters but if the model is to simple like `price ~ category.secondary` you will not be able to obtain good predictions. Choose the level of complexity that you consider appropriate to help the Research and Development (R & D) division and to determine the new price of the new product

**Note 2**: If you find that the model generates silly results, like for example negative or 0 values for the price, abandon the choosen model and use another model or change the product gap combination or discard particular locations where the product will going to be offer.

#### Fit the model

Pending

```{r}
#| label: tbl-model-2-tidy
#| tbl-cap: Model 2 (price = f(category secondary, frame))
model2 <- lm(
  formula = price ~ category.secondary + frame,
  data = bike_sales
)

model2_tidy <- model2 |> 
  tidy()

model2_tidy |> 
  datatable(
    colnames = c(
      "Term",
      "Estimated parameters",
      "Standard error",
      "Statistic",
      "P-value"
    )
  ) |> 
  formatRound(
    columns = c(2, 3, 4, 5),
    digits = 3
  )
```

#### Predict the price of the new product

Pending

```{r}
#| label: tbl-model2-tidy-predicted-price
#| tbl-cap: Price prediction based on category secondary and frame using the product gap
new_data <- tibble(
  category.secondary = "Fat Bike",
  frame = "Carbon"
)

model2_new_data_pred <- predict(
  object = model2,
  newdata = new_data
) |> 
  enframe(
    name = "Observation",
    value = "Predicted Price"
  ) |> 
  bind_cols(new_data)

model2_new_data_pred |> 
  datatable(
    colnames = c(
      "Observation",
      "Predicted price",
      "Category secondary",
      "Frame"
    )
  ) |> 
  formatRound(
    columns = c(2),
    digits = 2
  )
```

## Segment bike shops with k-means clustering

Pending

### Bike shop revenue trends

Pending

```{r}
#| message: false
bike_sales_total_revenue <- bike_sales |> 
  select(
    bikeshop.name,
    price.ext,
    model,
    category.primary,
    category.secondary,
    frame
  ) |> 
  group_by(
    bikeshop.name,
    model,
    category.primary,
    category.secondary,
    frame
  ) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  ungroup()
```

### Prepare data

Pending

#### Normalization

```{r}
bike_sales_total_revenue_pct <- bike_sales_total_revenue |> 
  group_by(bikeshop.name) |> 
  mutate(total_revenue_pct = total_revenue / sum(total_revenue)) |> 
  ungroup()
```

#### Customer-product table

Pending

```{r}
customer_product_table <- bike_sales_total_revenue_pct |> 
  select(
    bikeshop.name,
    model,
    total_revenue_pct
  ) |> 
  pivot_wider(
    id_cols = bikeshop.name,
    names_from = model,
    values_from = total_revenue_pct,
    values_fill = 0 
  )
```

```{r}
#| label: fig-customer-product-table
#| fig-cap: Graphical representation of the customer product table
#| fig-height: 12
customer_product_table_longer <- customer_product_table |> 
  pivot_longer(
    cols = -bikeshop.name,
    names_to = "model",
    values_to = "total_revenue_pct"
  )

tidyheatmap(
  df = customer_product_table_longer,
  rows = model,
  columns = bikeshop.name,
  values = total_revenue_pct,
  border_color = "black",
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  clustering_method = "complete"
)
```

### Performing k-means clustering

Pending

#### Choosing the number of clusters

```{r}
#| label: tbl-n-clust
#| tbl-cap: Choosing the number of clusters to use in k-means
#| warning: false
n_clust_customer_product_table <- customer_product_table |> 
  select(-bikeshop.name) |> 
  n_clusters(
    standardize = FALSE, # The data is using the same scale
    nbclust_method = "kmeans",
    n_max = 10
  )

n_clusters <- 4

n_clust_customer_product_table |> 
  as_tibble() |> 
  mutate(n_Clusters = as.integer(n_Clusters)) |> 
  datatable(
    colnames = c(
      "Number of Clusters",
      "Method",
      "R package",
      "Duration (seconds)"
    )
  ) |> 
  formatRound(
    columns = c(4),
    digits = 2
  )
```

### Applying k-means

Pending

```{r}
set.seed(seed = 1234)
kmeans_object <- customer_product_table |> 
  select(-bikeshop.name) |> 
  kmeans(
    centers = n_clusters, 
    algorithm = "Hartigan-Wong"  
  )
```

#### Extracting the cluster assignment

Pending

```{r}
#| label: tbl-clusters
#| tbl-cap: Cluster assignment using k-means
clusters <- kmeans_object |> 
  augment(
    data = customer_product_table
  ) |> 
  select(
    bikeshop.name,
    .cluster
  )

clusters |> 
  arrange(.cluster) |> 
  datatable(
    colnames = c(
      "Bikeshop name",
      "Cluster"
    )
  )
```

### Data visualization with k-means clustering and principal component analysis

Pending

```{r}
pca_object <- customer_product_table |> 
  select(-bikeshop.name) |> 
  prcomp(
    center = TRUE,
    scale. = FALSE
  )
```

#### Extracting the first 2 principal components

Pending

```{r}
pca_1_2 <- pca_object$x |> 
  as_tibble() |> 
  select(
    PC1,
    PC2
  )
```

#### Cluster visualization using k-means and principal component analysis

Using the information about the cluster assignment and the 2 principal components consolidate these parts into a tibble to generate a plot and visualize the clusters using the 2 dimensional projection

```{r}
#| label: fig-clusters-pca-1-2
#| fig-cap: Cluster assignment visualization using k-means and principal component analysis
clusters_pca_1_2 <- pca_1_2 |> 
  bind_cols(clusters)

clusters_pca_1_2 |> 
  ggplot() + 
  geom_point(
    aes(
      x = PC1, y = PC2,
      color = .cluster
    )
  ) +
  labs(
    x = "First principal component (18.9 %)",
    y = "Second principal component (11.8 %)"
  )
```

# Ai use disclaimer

## Tools

- DeepSeek
- Claude

## Uses

- [ ] Assistance in text writing
- [x] Assistance in image creation
- [x] Assistance in developing ideas and exploring arguments
- [x] Assistance with computer programming