---
title: "Marketing report part 3"
date: 2026-06-06
author: 
  - Norah Jones   
  - Bill Gates
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| echo: false
#| warning: false

library(sweep)
library(tidyverse)
library(scales)
library(tidymodels)
library(DT)
library(tidyheatmaps)
```

# Modeling

## Introduction

Pending

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## The product gap

Pending

### Prepare data

Pending

```{r}
#| message: false
total_sales_by__model_cat2_frame <- bike_sales |> 
  select(
    model,
    category.secondary,
    frame,
    price.ext
  ) |> 
  group_by(
    model, category.secondary, frame
  ) |> 
  summarise(
    total_revenue = sum(price.ext)
  ) |> 
  ungroup() |> 
  arrange(desc(total_revenue)) |> 
  mutate(category.secondary = fct_reorder(
    .f = category.secondary,
    .x = total_revenue
  )
)
```

### Visualization

Pending

```{r}
#| label: fig-total-sales-by-model-cat2-frame
#| fig-cap: Product gap and total sales by model, category secondary and frame
total_sales_by__model_cat2_frame |> 
  ggplot() + 
  geom_point(
    aes(x = frame, y = total_revenue),
    position = position_jitter(
      width = 0.2,
      height = 0.2
    )
  ) +
  scale_y_continuous(
    labels = label_currency()
  ) +
  facet_wrap(
    facets = vars(category.secondary)
  ) +
  labs(
    x = "Frame",
    y = "Total revenue (US dollars)"
  )
```

## Determine the price of a new product

Pending

### Selecting a new potential product

Pending

-   Cyclocross Aluminum
-   Fat Bike Carbon
-   Over Mountain Aluminum
-   Sport Carbon
-   Triathalon Aluminum

**Note**: choose 1 product from the product gap

### Building a model to price a new product

Pending

#### Choosing a model

Pending

-   `price ~ category.secondary + frame`

**Note 1**: If the model is to complex like `price ~ category.secondary*frame*bikeshop.state` you will obtain a considerable number of parameters but if the model is to simple like `price ~ category.secondary` you will not be able to obtain good predictions. Choose the level of complexity that you consider appropriate to help the Research and Development (R & D) division and to determine the new price of the new product

**Note 2**: If you find that the model generates silly results, like for example negative or 0 values for the price, abandon the choosen model and use another model or change the product gap combination or discard particular locations where the product will going to be offer.

#### Fit the model

Pending

```{r}
#| label: tbl-model-2-tidy
#| tbl-cap: Model 2 (price = f(category secondary, frame))
model2 <- lm(
  formula = price ~ category.secondary + frame,
  data = bike_sales
)

model2_tidy <- model2 |> 
  tidy()

model2_tidy |> 
  datatable(
    colnames = c(
      "Term",
      "Estimated parameters",
      "Standard error",
      "Statistic",
      "P-value"
    )
  ) |> 
  formatRound(
    columns = c(2, 3, 4, 5),
    digits = 3
  )
```

#### Predict the price of the new product

Pending

```{r}
#| label: tbl-model2-tidy-predicted-price
#| tbl-cap: Price prediction based on category secondary and frame using the product gap
new_data <- tibble(
  category.secondary = "Fat Bike",
  frame = "Carbon"
)

model2_new_data_pred <- predict(
  object = model2,
  newdata = new_data
) |> 
  enframe(
    name = "Observation",
    value = "Predicted Price"
  ) |> 
  bind_cols(new_data)

model2_new_data_pred |> 
  datatable(
    colnames = c(
      "Observation",
      "Predicted price",
      "Category secondary",
      "Frame"
    )
  ) |> 
  formatRound(
    columns = c(2),
    digits = 2
  )
```

## Segment bike shops with k-means clustering

Pending

### Bike shop revenue trends

Pending

```{r}
#| message: false
bike_sales_total_revenue <- bike_sales |> 
  select(
    bikeshop.name,
    price.ext,
    model,
    category.primary,
    category.secondary,
    frame
  ) |> 
  group_by(
    bikeshop.name,
    model,
    category.primary,
    category.secondary,
    frame
  ) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  ungroup()
```

### Prepare data

Pending

#### Normalization

```{r}
bike_sales_total_revenue_pct <- bike_sales_total_revenue |> 
  group_by(bikeshop.name) |> 
  mutate(total_revenue_pct = total_revenue / sum(total_revenue)) |> 
  ungroup()
```

#### Customer-product table

Pending

```{r}
customer_product_table <- bike_sales_total_revenue_pct |> 
  select(
    bikeshop.name,
    model,
    total_revenue_pct
  ) |> 
  pivot_wider(
    id_cols = bikeshop.name,
    names_from = model,
    values_from = total_revenue_pct,
    values_fill = 0 
  )
```

```{r}
#| label: fig-customer-product-table
#| fig-cap: Graphical representation of the customer product table
#| fig-height: 12
customer_product_table_longer <- customer_product_table |> 
  pivot_longer(
    cols = -bikeshop.name,
    names_to = "model",
    values_to = "total_revenue_pct"
  )

tidyheatmap(
  df = customer_product_table_longer,
  rows = model,
  columns = bikeshop.name,
  values = total_revenue_pct,
  border_color = "black",
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  clustering_method = "complete"
)
```

### Performing k-means clustering

The objective of this section is to apply the k-means clustering technique by grouping the bike shops

-   First using the function `n_clusters` from the package `parameters` on the customer-product table to identify the number of clusters to use for k-means

```{r}
# Follow this template to identify the number of clusters to use in k-means

# library(parameters)

# n_clusters_customer_product_table <- customer_product_table |>
#   select(-bikeshop.name) |>
#   n_clusters(standardize = FALSE,
#              nbclust_method = "kmeans")

# n_clusters_customer_product_table

# If you want, you can check the
# criteria or methods used to select
# the number of clusters to use in kmeans
# This part is optional

# n_clusters_customer_product_table |>
#   as_tibble()
```


Use the `kmeans` function with the argument `centers` equal to the number of clusters identify above and applying it to the customer product table

-   Remember to use `set.seed` to fix the initial centers. For example, you can use `set.seed(seed = 1234)`  

-   Remember that `kmeans` only works with numerical data so you will need to drop the `bikeshop.name` from the customer product table

    -   Checkout `?kmeans` and the argument `x` in relation to this aspect

-   Once you have apply the k-means clustering technique save this object as `kmeans_object`

```{r}
# Follow this template to apply k-means
# to the customer-product table

# set.seed(seed = 1234)
# kmeans_object <- customer_product_table |>
#   select(-bikeshop.name) |>
#   kmeans(centers = 4)
```

#### Extracting the cluster assignment

Now that you have apply the k-means clustering technique you are going to extract the clusters and assign them to each bike shop

You are going to use the `glance` function from the `tidymodels` library to extract the cluster assignment and select the bike shop names and the clusters to obtain the information required by the marketing team

```{r}
# Follow this template to obtain the relevant information
# required by the marketing team

# clusters <- kmeans_object |>
#   augment(data = customer_product_table) |>
#   select(bikeshop.name, .cluster)

## You can check the data using
### clusters |> glimpse()
### clusters |> count(.cluster)
```

### Data visualization with k-means clustering and principal component analysis

Pending

```{r}
pca_object <- customer_product_table |> 
  select(-bikeshop.name) |> 
  prcomp(
    center = TRUE,
    scale. = FALSE
  )
```

#### Extracting the first 2 principal components

Pending

```{r}
pca_1_2 <- pca_object$x |> 
  as_tibble() |> 
  select(
    PC1,
    PC2
  )
```

#### Cluster visualization using k-means and principal component analysis

Using the information about the cluster assignment and the 2 principal components consolidate these parts into a tibble to generate a plot and visualize the clusters using the 2 dimensional projection

```{r}
# Follow this template to visualize the fill clusters

# clusters_pca_1_2 <- pca_1_2 |>
#   bind_cols(clusters)

# clusters_pca_1_2 |>
#   ggplot() +
#   geom_point(aes(x = PC1,
#                  y = PC2,
#                  color = .cluster))

## Improve the plot by adding labels to the x, y axis
## and a title to explain the visualization to the
## marketing team
```

# Ai use disclaimer

## Tools

- DeepSeek
- Claude

## Uses

- [ ] Assistance in text writing
- [x] Assistance in image creation
- [x] Assistance in developing ideas and exploring arguments
- [x] Assistance with computer programming