---
title: "Relationships Between Continuous Variables"
author: 
  - Luis Francisco Gomez
  - Juan Perez
format: 
  html:
    embed-resources: true
    toc: true
    code-fold: true
    number-sections: true
editor: visual
---

```{r}
#| label: libraries
#| message: false

library(tidyverse)
library(skimr)
library(corrr)
library(DT)
```

# Import data

```{r}
#| message: false

customer <- read_csv(file = "../000_data/004_data_chapter4.csv")
```

# Inspect data

```{r}
customer |> glimpse()
```

# Transform data

```{r}
customer <- customer |>
  mutate(cust.id = factor(x = cust.id,
                          ordered = FALSE),
         email = factor(x = email,
                        ordered = FALSE),
         online.visits = as.integer(x = online.visits),
         online.trans = as.integer(x = online.trans),
         store.trans = as.integer(x = store.trans),
         sat.service = factor(x = sat.service,
                              ordered = TRUE),
         sat.selection = factor(x = sat.selection,
                                ordered = TRUE))
```

# Summarize data

```{r}
customer |> skim()
```

# Visualize data

## Physical stores vs online: spending

```{r}
customer |> 
  ggplot() +
  geom_point(aes(x = store.spend,
                 y = online.spend,
                 color = email)) +
  scale_color_manual(values = c("#ea5545",
                                "#b33dc6"), 
                     labels = c("Email in file: no",
                                "Email in file: yes")) +
  scale_x_continuous(transform = "log1p", 
                     breaks = c(1, 2, 5, 10, 20, 50, 100, 500)) +
  scale_y_continuous(transform = "log1p",
                     breaks = c(1, 5, 50, 500, 5000)) +
  labs(x = "Sales in the physical store (US dollars)",
       y = "Sales online store (US dollars)",
       title = "Yearly sales: physical store vs online",
       color = "Email")
```

# Summary statistics between continuous variables

## Pearson correlation

```{r}
coef_pearson <- cor(x = customer$store.spend, 
                    y = customer$online.spend,
                    method = "pearson")

coef_pearson
```

## Correlation matrices

```{r}
#| message: false

correlation_matrix <- customer |> 
  select(where(is.numeric)) |> 
  correlate(use = "pairwise.complete.obs",
            method = "pearson",
            diagonal = 1)

correlation_matrix |> 
  datatable()
```

### Visualizing correlation matrix

```{r}
correlation_matrix |> 
  autoplot(triangular = "lower")
```

## Transforming and visualizing data

### Preparing data

```{r}
customer <- customer |> 
  mutate(distance.to.store_trans = 1/distance.to.store,
         distance.to.store_log = log(distance.to.store))
```

### Visualizing transformed data

```{r}
customer |> 
  ggplot() + 
  geom_point(aes(x = distance.to.store_log,
                 y = store.spend))
```

### Calculating correlation

```{r}
cor(customer$distance.to.store_log,
    customer$store.spend)
```

## visualizing categorical values

### Using a scatterplot

```{r}
customer |> 
  ggplot() + 
  geom_point(aes(x = sat.service,
                 y = sat.selection),
             position = position_jitter(width = 0.2,
                                        height = 0.2))
```

### Using text

#### Prepare data

```{r}
customer_satisfaction <- customer |> 
  count(sat.service, sat.selection)
```

#### Creating the plot

```{r}
customer_satisfaction |> 
  ggplot() +
  geom_label(aes(x = sat.service,
                 y = sat.selection,
                 label = n))
```
