---
title: "An Overview of the R Language"
author: 
  - name: Luis Francisco Gomez Lopez
date: 2024-04-10
format: 
  html:
    embed-resources: true
    code-fold: true
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(skimr)
```

# Import data

This data correspond to a satisfaction survey about the experience of buying some product and also the rating about the person who sell the product

```{r}
#| message: false
satisfaction_data <- read_csv(file = "../000_data/002_data_chapter2.csv")
```

# Transforming data

`Segment` is a nominal variable taking into account that it doesn't have an order

```{r}
satisfaction_data <- satisfaction_data |> 
  mutate(Segment = factor(Segment, ordered = FALSE))
```

# Exploring data

We can use the `skim` function to summarize the data

```{r}
satisfaction_data |> skim()
```

# Creating new variables

Create an ID for the customers

```{r}
satisfaction_data <- satisfaction_data |> 
  mutate(Customer = 1:500)
```

# Selecting columns

```{r}
satisfaction_data |> 
  select(Segment, Customer)
```

# Filtering columns

```{r}
satisfaction_data |> 
  filter(iProdSAT >= 5)
```

# Calculating summary statistics

```{r}
satisfaction_data |> 
  summarise(mean_iSalesSAT = mean(iSalesSAT),
            mean_iSalesREC = mean(iSalesREC),
            max_iSalesSAT = max(iSalesSAT))
```

# Grouping variables

```{r}
satisfaction_data |> 
  group_by(Segment) |> 
  summarise(mean_iProdREC = mean(iProdREC))
```

# Manipulating data

```{r}
satisfaction_data |> 
  select(Customer, Segment, iProdSAT) |> 
  filter(iProdSAT >=5) |> 
  group_by(Segment) |> 
  summarise(mean_iProdSAT = mean(iProdSAT))
```
