---
title: "Marketing report part 2"
date: 2026-05-16
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gomez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| echo: false
#| warning: false

library(sweep)
library(tidyverse)
library(scales)
library(DT)
```

# Exploring data, visualizing it and applying statistical tests

## Introduction

Pending. Bla bla bla @fig-frame-category-secondary

Bla bla bla bla @tbl-revenue-stats-by-cat2-frame

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## Relation between continuous variables

Pending. Bla bla bla @fig-price-quantity-by-frame

### Prices vs quantity by frame

Pending

```{r}
#| label: fig-price-quantity-by-frame
#| fig-cap: Relationship between price and quantity by frame
cor_price_quantity <- cor(x = bike_sales$price, y = bike_sales$quantity) |> 
  format(scientific = FALSE)

bike_sales |> 
  ggplot() +
  geom_point(
    aes(x = price, y = quantity,
        color = frame)
  ) +
  scale_x_continuous(
    labels = label_currency()
  ) +
  labs(
    x = "Price (US Dollars/Frame)",
    y = "Quantity (Units)",
    color = "Frame type"
  )
```

There is not a linear relation between `price` and `quantity` because using the pearson correlation coefficient we obtain `{r} cor_price_quantity`

### Prices vs quantity transformed by frame

Pending

```{r}
#| label: fig-price-quantity-log10-trans-by-frame
#| fig-cap: Relationship between price and quantity transformed using a logarithmic base 10 scale by frame

cor_price_quantity_trans <- cor(
  x = log(bike_sales$price, base = 10), 
  y = log(bike_sales$quantity, base = 10)) |> 
  format(scientific = FALSE)

bike_sales |> 
  ggplot() +
  geom_point(
    aes(x = price, y = quantity,
        color = frame)
  ) +
  scale_x_continuous(
    labels = label_currency(accuracy = 1),
    transform = "log10"
  ) +
  scale_y_continuous(
    transform = "log10"
  ) +
  labs(
    x = "Price (US Dollars/Frame)",
    y = "Quantity (Units)",
    color = "Frame type"
  )
```

Even when applying a logarithmic transformation in base 10 we obtain the following pearson correlation coefficient `{r} cor_price_quantity_trans` which is practically 0.

## Relation between categorical variables

Pending

```{r}
#| label: fig-frame-category-secondary
#| fig-cap: Relationship between frame vs category secondary and the product gap
frame_category_secondary_count <- bike_sales |> 
  count(frame, category.secondary)

frame_category_secondary_count |> 
  ggplot() +
  geom_label(
    aes(
      x = frame,
      y = category.secondary,
      label = n
    )
  ) +
  labs(
    x = "Frame type",
    y = "Category Secondary"
  )
```

## Descriptive statistics by group

Pending

### Mean, median and standard deviation for revenue by `(category.secondary, frame)`

Pending

```{r}
#| message: false
#| label: tbl-revenue-stats-by-cat2-frame
#| tbl-cap: Revenue statistics by category secondary and frame

revenue_stats_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(
    mean_revenue   = mean(price.ext),
    median_revenue = median(price.ext),
    sd_revenue     = sd(price.ext) 
  ) |> 
  ungroup() |> 
  arrange(desc(mean_revenue))

revenue_stats_by_cat2_frame |> 
  datatable(
    colnames = c(
      "Category secondary",
      "Frame",
      "Mean",
      "Median",
      "Standard deviation"
    )
  ) |> 
  formatRound(
    columns = c(3, 5),
    digits = 2
  )
```

### Percentage of total revenue generated by `(category.secondary, frame)`

Pending

```{r}
#| message: false
#| label: tbl-pct-revenue-by-cat2-frame
#| tbl-cap: Total revenue and percentage of total revenue by category secondary and frame
pct_revenue_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup() |> 
  mutate(
    pct_revenue = (total_revenue / sum(total_revenue))*100
  ) |> 
  arrange(desc(pct_revenue))

pct_revenue_by_cat2_frame |> 
  datatable(
    colnames =  c(
      "Category secondary",
      "Frame",
      "Total revenue",
      "Percentage total revenue"
    )
  ) |> 
  formatRound(
    columns = 4
  )
```

### Mean, median and standard deviation for revenue by `bikeshop.name`

Pending

```{r}
#| message: false
#| label: tbl-revenue-stats-by-bikeshop
#| tbl-cap: Revenue statistics by bikeshop
revenue_stats_by_bikeshop <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(
    mean_revenue   = mean(price.ext),
    median_revenue = median(price.ext),
    sd_revenue     = sd(price.ext) 
  ) |> 
  ungroup() |> 
  arrange(desc(mean_revenue))

revenue_stats_by_bikeshop |> 
  datatable(
    colnames = c(
      "Bikeshop",
      "Mean",
      "Median",
      "Standard deviation"
    )
  ) |> 
  formatRound(
    columns = c(2, 4),
    digits = 2
  )
```

### Percentage of total revenue generated by each `bikeshop.name`

Pending

```{r}
pct_revenue_by_bikeshop <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup() |> 
  mutate(
    pct_revenue = (total_revenue / sum(total_revenue))*100
  ) |> 
  arrange(desc(pct_revenue))

pct_revenue_by_bikeshop |> 
  datatable(
    colnames =  c(
      "Bikeshop",
      "Total revenue",
      "Percentage total revenue"
    )
  ) |> 
  formatRound(
    columns = 3
  )
```

## Statistical tests

### Testing group frequencies: `chisq.test()`

Inspect, using an statistical test, if the total revenue is distributed in a homogeneous way by `category.secondary` and `frame` using a $\chi^2$ test for a given vector of probabilities.

Taking into account that there are 13 available models offer by the corporation use $p_i = \frac{1}{13}$ for $i = 1, 2, \ldots, 13$ in the $\chi^2$ test

Once you apply the test point out your conclusion taking into account the p-value of the test

```{r}
# To attack the problem you need to really understand
# the chi squared test for given probabilities

## First, prepare the data in a similar or in the following
## way
### table_category_secondary_frame <- bike_sales |>
###   group_by(category.secondary, frame) |>
###   summarise(total_revenue = sum(price.ext),
###             .groups = "drop") |>
###   mutate(total_revenue_pct = (total_revenue / sum(total_revenue))*100) |>
###   arrange(desc(total_revenue_pct))

## Second, create a new variable to identify
## the models in a similar or in the following way
### table_models <- table_category_secondary_frame |>
###  mutate(model = str_c(category.secondary, frame, sep = " ")) |>
###  select(model, total_revenue)
#### For more information about str_c check out the examples in
#### https://stringr.tidyverse.org/reference/str_c.html

## Third, create a named vector to apply the chi test
## taking into account what is mention in ?chisq.test
## in the section related to the Examples and the part
## about Testing for population probabilities
### To create a named vector check out the deframe
### function from the tidyverse
#### https://tibble.tidyverse.org/reference/enframe.html
##### deframe(table_models)

## Fourth, apply the chi squared test
### deframe(table_models) |>
###  chisq.test(p = rep.int(x = 1/13, times = 13))
#### rep.int(x = 1/13, times = 13) is just a way
#### to repeat 1/13 thirteen times
##### Check out ?rep.int to understand it

## Fifth, use the tidymodels library:
## library(tidymodels) using tidy
### Remember to load the library in the first code
### chunk: library(tidymodels)
#### For more information check out
#### https://broom.tidymodels.org/reference/tidy.htest.html
##### deframe(table_models) |>
#####  chisq.test(p = rep.int(x = 1/13, times = 13)) |>
#####  tidy()
```

Inspect, using an statistical test, if the total revenue is distributed in a homogeneous way by `bikeshop.name` using a $\chi^2$ test for a given vector of probabilities.

Taking into account that there are 30 bike shops offer use $p_i = \frac{1}{30}$ for $i = 1, 2, \ldots, 30$ in the $\chi^2$ test

Once you apply the test point out your conclusion taking into account the p-value of the test

```{r}
# Apply a similar process as in the case of
# applying the chi squared test for
# category.secondary and frame in relation to
# the total revenue
## See the comments in that section for more
## information about how to apply the test
```

### Testing group means: `t.test()`

Inspect, using a two sample t-tests, if the mean revenue between the `Aluminum` and `Carbon` bicycle models are the same in relation to the `frame` variable

Once you apply the test point out your conclusion taking into account the p-value of the test

```{r}
# If you use base R you can apply t.test
## Check out more information in ?t.test
###
# If you use the tidyverse framework you can
# use t_test from tidymodels
## For more information check out
### https://infer.tidymodels.org/reference/t_test.html
```

### Analysis of Variance (ANOVA)

Apply an analysis of variance (ANOVA) using `price` as a response variable

-   First begin with a simple model

    -   `price ~ category.secondary`

-   Then add another variable

    -   `price ~ category.secondary + frame`

-   Expand the model in the following way where you can use the following equivalent expressions

    -   `price ~ category.secondary + frame + category.secondary:frame`
    -   `price ~ category.secondary*frame`[^1]

[^1]: `category.secondary*frame` is a short way to specify all the possible interactions between these 2 categorical variables which is equivalent to `category.secondary + frame + category.secondary:frame`

```{r}
# Use the functions aov and anova from base R
# to apply the analysis of variance (ANOVA)
## As an example you can begin with a simple model
### aov(formula = price ~ category.secondary,
###    data = bike_sales) |>
###  anova() |>
###  tidy()

## Then change the formula argument with the
## different alternative models
```

Taking into account the analysis of variance, point out if it make sense to add `category.secondary`, `frame` and its interaction to the model.

Depending on the answer to the above question point out if it make sense to try the following model:

-   `price ~ category.secondary*frame*bikeshop.state`[^2]

[^2]: `category.secondary*frame*bikeshop.state` is a way to specify all the possible interactions between these 3 categorical variables. In this case the model will have $2^3 - 1 = 7$ terms

# Ai use disclaimer

The following generative artificial intelligence tools where used:

- Gemini
- Claude

These tools were used in the following way:

- [ ] Assistance in text writing
- [x] Assistance in image creation
- [ ] Assistance in developing ideas and exploring arguments
- [x] Assistance with computer programming
