---
title: "Marketing report part 2"
date: 2026-05-16
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gomez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| echo: false
#| warning: false

library(sweep)
library(tidyverse)
library(scales)
library(DT)
library(tidymodels)
```

# Exploring data, visualizing it and applying statistical tests

## Introduction

Pending. Bla bla bla @fig-frame-category-secondary

Bla bla bla bla @tbl-revenue-stats-by-cat2-frame

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## Relation between continuous variables

Pending. Bla bla bla @fig-price-quantity-by-frame

### Prices vs quantity by frame

Pending

```{r}
#| label: fig-price-quantity-by-frame
#| fig-cap: Relationship between price and quantity by frame
cor_price_quantity <- cor(x = bike_sales$price, y = bike_sales$quantity) |> 
  format(scientific = FALSE)

bike_sales |> 
  ggplot() +
  geom_point(
    aes(x = price, y = quantity,
        color = frame)
  ) +
  scale_x_continuous(
    labels = label_currency()
  ) +
  labs(
    x = "Price (US Dollars/Frame)",
    y = "Quantity (Units)",
    color = "Frame type"
  )
```

There is not a linear relation between `price` and `quantity` because using the pearson correlation coefficient we obtain `{r} cor_price_quantity`

### Prices vs quantity transformed by frame

Pending

```{r}
#| label: fig-price-quantity-log10-trans-by-frame
#| fig-cap: Relationship between price and quantity transformed using a logarithmic base 10 scale by frame

cor_price_quantity_trans <- cor(
  x = log(bike_sales$price, base = 10), 
  y = log(bike_sales$quantity, base = 10)) |> 
  format(scientific = FALSE)

bike_sales |> 
  ggplot() +
  geom_point(
    aes(x = price, y = quantity,
        color = frame)
  ) +
  scale_x_continuous(
    labels = label_currency(accuracy = 1),
    transform = "log10"
  ) +
  scale_y_continuous(
    transform = "log10"
  ) +
  labs(
    x = "Price (US Dollars/Frame)",
    y = "Quantity (Units)",
    color = "Frame type"
  )
```

Even when applying a logarithmic transformation in base 10 we obtain the following pearson correlation coefficient `{r} cor_price_quantity_trans` which is practically 0.

## Relation between categorical variables

Pending

```{r}
#| label: fig-frame-category-secondary
#| fig-cap: Relationship between frame vs category secondary and the product gap
frame_category_secondary_count <- bike_sales |> 
  count(frame, category.secondary)

frame_category_secondary_count |> 
  ggplot() +
  geom_label(
    aes(
      x = frame,
      y = category.secondary,
      label = n
    )
  ) +
  labs(
    x = "Frame type",
    y = "Category Secondary"
  )
```

## Descriptive statistics by group

Pending

### Mean, median and standard deviation for revenue by `(category.secondary, frame)`

Pending

```{r}
#| message: false
#| label: tbl-revenue-stats-by-cat2-frame
#| tbl-cap: Revenue statistics by category secondary and frame

revenue_stats_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(
    mean_revenue   = mean(price.ext),
    median_revenue = median(price.ext),
    sd_revenue     = sd(price.ext) 
  ) |> 
  ungroup() |> 
  arrange(desc(mean_revenue))

revenue_stats_by_cat2_frame |> 
  datatable(
    colnames = c(
      "Category secondary",
      "Frame",
      "Mean",
      "Median",
      "Standard deviation"
    )
  ) |> 
  formatRound(
    columns = c(3, 5),
    digits = 2
  )
```

### Percentage of total revenue generated by `(category.secondary, frame)`

Pending

```{r}
#| message: false
#| label: tbl-pct-revenue-by-cat2-frame
#| tbl-cap: Total revenue and percentage of total revenue by category secondary and frame
pct_revenue_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup() |> 
  mutate(
    pct_revenue = (total_revenue / sum(total_revenue))*100
  ) |> 
  arrange(desc(pct_revenue))

pct_revenue_by_cat2_frame |> 
  datatable(
    colnames =  c(
      "Category secondary",
      "Frame",
      "Total revenue",
      "Percentage total revenue"
    )
  ) |> 
  formatRound(
    columns = 4
  )
```

### Mean, median and standard deviation for revenue by `bikeshop.name`

Pending

```{r}
#| message: false
#| label: tbl-revenue-stats-by-bikeshop
#| tbl-cap: Revenue statistics by bikeshop
revenue_stats_by_bikeshop <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(
    mean_revenue   = mean(price.ext),
    median_revenue = median(price.ext),
    sd_revenue     = sd(price.ext) 
  ) |> 
  ungroup() |> 
  arrange(desc(mean_revenue))

revenue_stats_by_bikeshop |> 
  datatable(
    colnames = c(
      "Bikeshop",
      "Mean",
      "Median",
      "Standard deviation"
    )
  ) |> 
  formatRound(
    columns = c(2, 4),
    digits = 2
  )
```

### Percentage of total revenue generated by each `bikeshop.name`

Pending

```{r}
pct_revenue_by_bikeshop <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup() |> 
  mutate(
    pct_revenue = (total_revenue / sum(total_revenue))*100
  ) |> 
  arrange(desc(pct_revenue))

pct_revenue_by_bikeshop |> 
  datatable(
    colnames =  c(
      "Bikeshop",
      "Total revenue",
      "Percentage total revenue"
    )
  ) |> 
  formatRound(
    columns = 3
  )
```

## Statistical tests

Pending

### Testing group frequencies: `chisq.test()`

Pending

#### `chisq.test()`: Total revenue by category secondary and frame

Pending

```{r}
#| message: false
#| label: tbl-chi-test_total-revenue-cat2-frame
#| tbl-cap: Chi square test for given probabilities for total revenue by category secondary and frame 
total_revenue_by_cat2_frame <- bike_sales |> 
  group_by(category.secondary, frame) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup() 

table_models <- total_revenue_by_cat2_frame |> 
  mutate(model = str_c(
    category.secondary, frame, sep = " "
  )
) |> 
select(model, total_revenue)

chi_test_total_revenue_cat2_frame <- table_models |> 
  deframe() |> 
  chisq.test(
    p = rep.int(x = 1 / 13, times = 13) 
  ) |> 
  tidy()

chi_test_total_revenue_cat2_frame |> 
  datatable(
    colnames = c(
      "Statistic",
      "P-value",
      "Degrees of freedom",
      "Method"
    )
  ) |> 
  formatRound(
    columns = c(1), 
    digits = 0
  )
```

#### `chisq.test()`: Total revenue by bikeshop

Pending

```{r}
#| label: tbl-chi-test-total-revenue-bikeshop
#| tbl-cap: Chi squared test for given probabilities for total revenue by bikeshops 
total_revenue_by_bikeshop <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(
    total_revenue = sum(price.ext), 
  ) |> 
  ungroup()

chi_test_total_revenue_bikeshop <- total_revenue_by_bikeshop |> 
  deframe() |> 
  chisq.test(
    p = rep.int(x = 1 / 30, times = 30) 
  ) |> 
  tidy()

chi_test_total_revenue_bikeshop |> 
  datatable(
        colnames = c(
      "Statistic",
      "P-value",
      "Degrees of freedom",
      "Method"
    )
  ) |> 
  formatRound(
    columns = c(1),
    digits = 0
  )
```

### Testing group means: `t.test()`

Pending

```{r}
#| label: tbl-t-test-mean-revenue-by-frame
#| tbl-cap: T-test for mean revenue by frame
alpha <- 0.05

t_test_mean_revenue_by_frame <- bike_sales |> 
  t_test(
    formula = price.ext ~ frame,
    order = c("Aluminum", "Carbon"),
    alternative = "two-sided",
    mu = 0,
    conf_level = 1 - alpha
  )

t_test_mean_revenue_by_frame |> 
  select(-c(lower_ci, upper_ci)) |> 
  datatable(
    colnames = c(
      "Statistic",
      "Degrees of freedom",
      "P-value",
      "Alternative hypothesis",
      "Mean difference estimation"
    )
  ) |> 
  formatRound(
    columns = c(1, 2, 3, 5),
    digits = 3
  )
```

### Analysis of Variance (ANOVA)

Pending

-   `price ~ category.secondary + frame + category.secondary:frame`

Bla bla bla

```{r}
#| label: tbl-anova-model-3
#| tbl-cap: Analysis of variance (anova) to explain price based on category secondary, frame an their interactions 
model3 <- aov(
  formula = price ~ category.secondary + frame + category.secondary:frame,
  data = bike_sales
  ) |> 
  anova() |> 
  tidy()

model3 |> 
  datatable(
    colnames = c(
      "Term",
      "Degrees of freedom",
      "Sum of squares",
      "Mean or squares",
      "Statistic",
      "P-value"
    )
  ) |> 
  formatRound(
    columns = c(3, 4, 5, 6),
    digits = 3
  )
```

Bla bla bla

# Ai use disclaimer

The following generative artificial intelligence tools where used:

- Gemini
- Claude

These tools were used in the following way:

- [ ] Assistance in text writing
- [x] Assistance in image creation
- [ ] Assistance in developing ideas and exploring arguments
- [x] Assistance with computer programming
