---
title: Relationships Between Continuous Variables
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: ../000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: true
    warning: false
    fig.align: center
    out.width: 85%
lang: en
bibliography: ../000_references/marketing_research_and_analytics.bib
csl: ../000_references/apa-7.csl
editor: source
---

```{r}
#| label: libraries
#| echo: false

library(tidyverse)
library(skimr)
```

```{r}
#| label: palette
#| echo: false

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   This presentation is based on [@chapman_r_2019, Chapter 4]

## 

- **Purpose**

    - Understand the relationships between pairs of variables in multivariate data and examine how to visualize the relationships and compute statistics that describe their associations

# CRM system data

## {.smaller}

-   **cust.id**: customer identifier
-   **age**: decimal age in years
-   **credit.score**: 3-digit number in \[300, 900\], representing the credit risk
-   **email**: whether or not there is information about the customer email
-   **distance.to.store**: distance in kilometers to the nearest physical store
-   **online.visits**: yearly visits to the online store
-   **online.trans**: yearly online orders
-   **online.spend**: yearly spending in those online orders
-   **store.trans**: yearly orders in physical stores
-   **store.spend**: yearly spending in those physical store orders

## {.smaller}

-   **sat.service**: satisfaction with service using an ordinal 5 point scale and collected using a survey

-   **sat.selection**: satisfaction with product selection using an ordinal 5 point scale and collected using a survey

    -   **Ordinal 5 point scale used and possible values in the survey**:

        -   Extremely satisfied: 5
        -   Very satisfied: 4
        -   Moderately satisfied: 3
        -   Very unsatisfied: 2
        -   Extremely unsatisfied: 1
        -   NA: customer did not response the survey

##

-   **Import data**

```{r}
customer <- read_csv(file = "http://goo.gl/PmPkaG")
customer |> head(n=5)
```

## {.smaller}

-   **Transform data**

```{r}
customer <- customer |>
  mutate(cust.id = factor(x = cust.id, ordered = FALSE),
         email = factor(x = email, ordered = FALSE),
         online.visits = as.integer(x = online.visits),
         online.trans = as.integer(x = online.trans),
         store.trans = as.integer(x = store.trans),
         sat.service = factor(x = sat.service, ordered = TRUE),
         sat.selection = factor(x = sat.selection, ordered = TRUE))
customer |> head(n=5)
```

##

-   **Inspect data**

```{r}
customer |> glimpse()
```

##

-   **Summarize data**

    -   Ups the table is really big!!! Try it in your console to see the complete table

```{r}
#| eval: false
customer |> skim()
```

##

-   **Scatterplots: the base R way**

```{r}
plot(x = customer$store.spend, y = customer$online.spend)
```

##

-   **Scatterplots: the base R way**

```{r}
plot(x = customer$store.spend, y = customer$online.spend,
     cex=0.7,
     main="Customers as of June 2014", 
     xlab="Prior 12 months in-store sales ($)", 
     ylab="Prior 12 months online sales ($)")
```

## {.smaller}

-   **Scatterplots: the base R way**

```{r}
my.col <- c("black", "green3")
my.pch <- c(1, 19)
plot(x = customer$store.spend, y = customer$online.spend,
     cex=0.7, col=my.col[customer$email], pch=my.pch[customer$email],
     main="Customers as of June 2014", 
     xlab="Prior 12 months in-store sales ($)", 
     ylab="Prior 12 months online sales ($)")
```

## {.smaller}

-   **Scatterplots: the base R way**

```{r}
my.col <- c("black", "green3")
my.pch <- c(1, 19)
plot(x = customer$store.spend, y = customer$online.spend,
     cex=0.7, col=my.col[customer$email], pch=my.pch[customer$email],
     main="Customers as of June 2014", 
     xlab="Prior 12 months in-store sales ($)", 
     ylab="Prior 12 months online sales ($)")
legend(x="topright", legend=paste("email on file:", levels(customer$email)), col=my.col, pch=my.pch)
```

## {.smaller}

-   **Scatterplots: the base R way**

```{r}
my.col <- c("black", "green3")
my.pch <- c(1, 19)
plot(x = customer$store.spend + 1, y = customer$online.spend + 1,
     cex=0.7, col=my.col[customer$email], pch=my.pch[customer$email],
     log ="xy",
     main="Customers as of June 2014", 
     xlab="Prior 12 months in-store sales ($)", 
     ylab="Prior 12 months online sales ($)")
legend(x="topright", legend=paste("email on file:", levels(customer$email)), col=my.col, pch=my.pch)
```

##

-   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 4

customer |> ggplot() +
  geom_point(aes(x = store.spend, y = online.spend, color = email))
```

## {.smaller}

-   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 4

customer |> ggplot() +
  geom_point(aes(x = store.spend, y = online.spend, color = email, shape = email)) +
  scale_color_manual(values = c("black", "green3"), labels = c("email on file: no", "email on file: yes")) +
  scale_shape_manual(values = c(1, 19), labels = c("email on file: no", "email on file: yes"))
```

## {.smaller}

-   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 4

customer |> ggplot() +
  geom_point(aes(x = store.spend, y = online.spend, color = email, shape = email)) +
  scale_color_manual(values = c("black", "green3"), labels = c("email on file: no", "email on file: yes")) +
  scale_shape_manual(values = c(1, 19), labels = c("email on file: no", "email on file: yes")) +
  scale_x_continuous(trans = "log1p", breaks = c(1, 2, 5, 10, 20, 50, 100, 500)) + 
  scale_y_continuous(trans = "log1p", breaks = c(1, 5, 50, 500))
```

## {.smaller}

-   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 3.5

customer |> ggplot() +
  geom_point(aes(x = store.spend, y = online.spend, color = email, shape = email)) +
  scale_color_manual(values = c("black", "green3"), labels = c("email on file: no", "email on file: yes")) +
  scale_shape_manual(values = c(1, 19), labels = c("email on file: no", "email on file: yes")) +
  scale_x_continuous(trans = "log1p", breaks = c(1, 2, 5, 10, 20, 50, 100, 500)) + 
  scale_y_continuous(trans = "log1p", breaks = c(1, 5, 50, 500)) +
  labs(x = "Prior 12 months in-store sales ($)", y = "Prior 12 months online sales ($)",
       title = "Customers as of June 2014")
```

##

-   **Correlation Coefficients**

    -   Pearson correlation coefficient for a sample

$$r_{xy} = \frac{\sum_{i=1}^n (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum_{i=1}^n (x_i - \bar{x})^2}\sqrt{\sum_{i=1}^n (y_i - \bar{y})^2}}$$ 

Where $n$ is the sample size, we must have paired numeric data $\{ (x_1, y_1), ..., (x_n, y_n) \}$, $\bar{x} = \sum_{i=1}^n x_i$ and $\bar{y} = \sum_{i=1}^n y_i$

-   This is a "nasty" formula but we can brake it down in smaller chunks

##

-   **Correlation Coefficients**

    -   Pearson correlation coefficient for a sample

```{r}
age_mean <- mean(customer$age)
age_credit.score <- mean(customer$credit.score)
numerator <- sum((customer$age - age_mean) * (customer$credit.score - age_credit.score)) 
denominator <- sqrt(sum((customer$age - age_mean)^2)) * sqrt(sum(((customer$credit.score - age_credit.score)^2)))
pearson_corr <- numerator / denominator
pearson_corr
```

-   **But don't worry be happy!!!: Use `cor`**

```{r}
cor(customer$age, customer$credit.score, method = 'pearson')
```

## {.smaller}

-   **Correlation matrices**

    -   Pearson correlation coefficients for samples in a tibble

```{r}
library(corrr) # Remember to install the package if it is not installed
correlation_matrix <- customer |> 
  select(where(is.numeric)) |> 
  correlate(use = "pairwise.complete.obs", # There are NA values
            method = "pearson",
            diagonal = NA)
correlation_matrix # Ups!!! The tibble is wide. Check out the tibble in your console
```

## {.smaller}

-   **Correlation matrices**

    -   Pearson correlation coefficients for samples in a tibble

```{r}
#| fig-height: 4

correlation_matrix |> autoplot(triangular = "lower")
```

## {.smaller}

-   **Transforming variables**

```{r}
cor(customer$store.spend, customer$distance.to.store)
```

```{r}
#| fig-height: 4

customer |> ggplot() +
  geom_point(aes(x = distance.to.store, y = store.spend))
```

## {.smaller}

-   **Transforming variables**

```{r}
cor(customer$store.spend, 1 / sqrt(customer$distance.to.store))
```

```{r}
#| fig-height: 4

customer |>
  mutate(distance.to.store_trans = 1 / sqrt(distance.to.store)) |>
  ggplot() +
  geom_point(aes(x = distance.to.store_trans, y = store.spend))
```

## {.smaller}

-   **Transforming variables**

    -   Understanding the logic behind inverse square root distance

```{r}
#| fig-height: 4

ggplot() +
  geom_function(fun = function(x) {1 / sqrt(x)}) +
  labs(x = "Distance",
       y = "Inverse square root distance")
```

## {.smaller}

-   **Visualizing categorical variables**

    -   **Scatterplots: the base R way**

```{r}
plot(as.integer(customer$sat.service), as.integer(customer$sat.selection),
     xlab = "Customer Satisfaction with Service",
     ylab = "Customer Satisfaction with Selection",
     main = "Customer as of June 2014")
```

##  {.smaller}

-   **Visualizing categorical variables**

    -   **Scatterplots: the base R way**

```{r}
plot(jitter(as.integer(customer$sat.service)), jitter(as.integer(customer$sat.selection)),
     xlab = "Customer Satisfaction with Service",
     ylab = "Customer Satisfaction with Selection",
     main = "Customer as of June 2014")
```

## {.smaller}

-   **Visualizing categorical variables**

    -   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 3.5

customer |> 
  ggplot() +
  geom_point(aes(x = sat.service, y = sat.selection)) +
  labs(x = "Customer Satisfaction with Service",
       y = "Customer Satisfaction with Selection",
       title = "Customer as of June 2014")
```

## {.smaller}

-   **Visualizing categorical variables**

    -   **Scatterplots: the tidyverse way**

```{r}
#| fig-height: 3.5

customer |> 
  ggplot() +
  geom_point(aes(x = sat.service, y = sat.selection), 
             position = position_jitter(width = 0.2, height = 0.2)) +
  labs(x = "Customer Satisfaction with Service",
       y = "Customer Satisfaction with Selection",
       title = "Customer as of June 2014")
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://posit.co/products/open-source/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**skimr**](https://CRAN.R-project.org/package=skimr), [**corrr**](https://CRAN.R-project.org/package=corrr) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}