---
title: "Comparing Groups: Tables and Visualizations"
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: ../000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: true
    warning: false
    fig.align: center
    out.width: 85%
lang: en
bibliography: ../000_references/marketing_research_and_analytics.bib
csl: ../000_references/apa-7.csl
editor: source
---

```{r}
#| label: libraries
#| echo: false

library(tidyverse)
```

```{r}
#| label: palette
#| echo: false

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   This presentation is based on [@chapman_r_2019, Chapter 5]

## 

- **Purpose**

    - Use descriptive summaries by groups and visualize them to investigate differences between groups

# Consumer segmentation survey

##

-   **age**: age of the consumer in years
-   **gender**: if the consumer is male of female
-   **income**: yearly disposable income of the consumer
-   **kids**: number of children of the consumer
-   **ownHome**: if the consumer owns a home
-   **subscribe**: if the consumer is subscribed or not
-   **Segment**: market segment assigned by a clustering algorithm [@chapman_r_2019, chapter 11], expert assignment or a segmentation typing tool

##

-   **Segment**:

    -   **Moving up**: consumers experiencing upward mobility in terms of their socioeconomic status
    -   **Suburb mix**: consumers living in suburban areas
    -   **Travelers**: consumers who prioritize experiences and adventures
    -   **Urban Hip**: consumers interested in urban culture, artistic expression, and modern trends

##

-   **Import data**

```{r}
segmentation <- read_csv(file = "http://goo.gl/qw303p")
segmentation |> head(n = 5)
```

##

-   **Transform data**

```{r}
segmentation <- segmentation |>
  mutate(gender = factor(gender, ordered = FALSE),
         kids = as.integer(kids),
         ownHome = factor(ownHome, ordered = FALSE),
         subscribe = factor(subscribe, ordered = FALSE),
         Segment = factor(Segment, ordered = FALSE))

segmentation |> head(n = 5)
```

## {.smaller}

-   **Basic Formula Syntax**

    -   $\sim$ and $+$: operators

    -   $y$: response variable

    -   $x, z$: explanatory variables

    -   $y \sim x + z$: a formula which means that $y$ depends on $x$ and $z$

        -   $+$ is used to indicate the addition of predictor variables to the right of the formula
        -   Be careful not to confuse the arithmetic operator $+$ with $+$ within a formula

```{r}
#| eval: false
?`+` # Arithmetic Operators
?formula # operators in a formula
```

##

-   **Descriptives for n-Way Groups: the base R way**

    -   Split data into $n$ subsets and compute summary statistics

```{r}
aggregate(x = income ~ Segment + ownHome, 
          data = segmentation, FUN = mean)
```

##

-   **Descriptives for n-Way Groups: the base R way**

    -   Split data into $n$ subsets and compute summary statistics

```{r}
aggregate(x = kids ~ Segment + ownHome, 
          data = segmentation, FUN = sum)
```

##

-   **Descriptives for n-Way Groups: the tidyverse way**

    -   Split data into $n$ subsets and compute summary statistics

```{r}
segmentation |> 
  group_by(Segment, ownHome) |>
  summarise(mean_income = mean(income))
```

##

-   **Descriptives for n-Way Groups: the tidyverse way**

    -   Split data into $n$ subsets and compute summary statistics

```{r}
segmentation |> 
  group_by(Segment, ownHome) |>
  summarise(sum_kids = sum(kids))
```

## {.smaller}

-   **Basic Formula Syntax**

    -   $\sim$, $+$ and $|$: operators

    -   $y$: response variable

    -   $x$: explanatory variable

    -   $z$: grouping variable

    -   $y \sim x | z$: $y$ depends on $x$ based on different groups defined by $z$

        -   $|$ is used to separate the grouping variable from the explanatory variable
        -   Be careful not to confuse the logical operator $|$ with $|$ within a formula

```{r}
#| eval: false
?`|` # Logical Operators
?lattice::xyplot # operators in a formula (you need first to install the package lattice)
```

## {.smaller}

-   **Visualization by group as frequencies: the lattice way**

```{r}
library(lattice)
histogram(~ subscribe | Segment + ownHome, data = segmentation)
```

## {.smaller}

-   **Visualization by group as frequencies: the tidyverse way**

```{r}
# Prepare data
subscriber_by_segment_home_ownership <- segmentation |>
  count(subscribe, Segment, ownHome) |>
  group_by(Segment, ownHome) |>
  mutate(n_pct = (n / sum(n)) * 100) |>
  ungroup() |> 
  arrange(Segment, ownHome)

subscriber_by_segment_home_ownership
```

## {.smaller}

-   **Visualization by group as frequencies: the tidyverse way**

```{r}
#| fig-height: 4

subscriber_by_segment_home_ownership |>
  ggplot() + 
  geom_col(aes(x = subscribe, y=n_pct)) + 
  facet_wrap(facets = vars(Segment,ownHome), 
             nrow = 2, ncol = 4)
```

## {.smaller}

-   **Visualization by group as proportions: the lattice way**

```{r}
# Prepare data
prop_table <- table(segmentation$subscribe,
                    segmentation$Segment) |>
  prop.table(margin = 2) |>
  _[2, ] # You can use _ as a placeholder. Check ?pipeOp
```

```{r}
#| fig-height: 4

barchart(prop_table, 
         xlab='Subscriber proportion by Segment', col='darkolivegreen')
```

## {.smaller}

-   **Visualization by group as proportions: the tidyverse way**

```{r}
# Prepare data
prop_table <- segmentation |>
  count(subscribe, Segment) |>
  group_by(Segment) |>
  mutate(n_pct = n / sum(n)) |>
  filter(subscribe == 'subYes')
```

```{r}
#| fig-height: 3

prop_table |> ggplot() + 
  geom_col(aes(x=n_pct, y=Segment),
           fill='darkolivegreen') +
  labs(x='Subscriber proportion by Segment', 
       y=NULL)
```

## {.smaller}

-   **Visualization by group with continuous data: the lattice way**

```{r}
# Prepare data
seg_income_agg <- aggregate(income ~ Segment + ownHome, 
                            data=segmentation, FUN = mean)
```

```{r}
barchart(income ~ Segment, data = seg_income_agg,
         groups=ownHome, auto.key=TRUE, # Add groups
         par.settings=simpleTheme(col=terrain.colors(n = 2))) # Change default colors
```

## {.smaller}

-   **Visualization by group with continuous data: the tidyverse way**

```{r}
# Prepare data
seg_income_agg <- segmentation |> 
  group_by(Segment, ownHome) |> 
  summarise(mean_income = mean(income)) |> 
  ungroup()
```

```{r}
#| fig-height: 3.5

seg_income_agg |> ggplot() +
  geom_col(aes(x=Segment, y=mean_income, fill=ownHome),
           position = position_dodge(), color='black') +
  scale_fill_manual(values=terrain.colors(n = 2))
```

## {.smaller}

-   **Visualization by group with continuous data: the lattice way**

```{r}
#| fig-height: 4.5

bwplot(Segment ~ income | ownHome,
       data = segmentation,
       xlab = 'Income')
```

## {.smaller}

-   **Visualization by group with continuous data: the tidyverse way**

```{r}
#| fig-height: 4

segmentation |> ggplot() +
  geom_boxplot(aes(x=income, y=Segment)) +
  facet_wrap(facets = vars(ownHome)) +
  labs(x='Income',
       y=NULL)
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://posit.co/products/open-source/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**lattice**](https://CRAN.R-project.org/package=lattice) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}