---
title: "Comparing Groups: Tables and Visualizations"
author: 
  - Luis Francisco Gómez López
institute: 
  - FAEDIS
date: last-modified
format: 
  beamer:
    colortheme: dolphin
    fonttheme: structurebold
    theme: AnnArbor
    link-citations: true
    colorlinks: true
    toc: true
    slide-level: 2
    section-titles: false
    include-in-header:
      file: ../000_tex/preamble.tex
fig-cap-location: bottom
tbl-cap-location: top
knitr: 
  opts_chunk: 
    echo: true
    warning: false
    fig-align: center
    out-width: 50%
lang: en
bibliography: ../000_references/marketing_research_and_analytics.bib
---

```{r}
#| echo: false
#| label: libraries

library(tidyverse)
```

# Please Read Me

##

-   This presentation is based on [@chapman_r_2019, Chapter 5]

# Purpose

##

-   Use descriptive summaries by groups and visualize them to investigate differences between groups

# Consumer segmentation survey

##

-   **age**: age of the consumer in years
-   **gender**: if the consumer is male of female
-   **income**: yearly disposable income of the consumer
-   **kids**: number of children of the consumer
-   **ownHome**: if the consumer owns a home
-   **subscribe**: if the consumer is subscribed or not
-   **Segment**: market segment assigned by a clustering algorithm [@chapman_r_2019, chapter 11], expert assignment or a segmentation typing tool

##

-   **Segment**:

    -   **Moving up**: consumers experiencing upward mobility in terms of their socioeconomic status
    -   **Suburb mix**: consumers living in suburban areas
    -   **Travelers**: consumers who prioritize experiences and adventures
    -   **Urban Hip**: consumers interested in urban culture, artistic expression, and modern trends

##

-   **Import data**

\tiny

```{r}
segmentation <- read_csv(file = "http://goo.gl/qw303p")
segmentation |> head(n = 5)
```

##

-   **Transform data**

\tiny

```{r}
segmentation <- segmentation |>
  mutate(gender = factor(gender, ordered = FALSE),
         kids = as.integer(kids),
         ownHome = factor(ownHome, ordered = FALSE),
         subscribe = factor(subscribe, ordered = FALSE),
         Segment = factor(Segment, ordered = FALSE))

segmentation |> head(n = 5)
```

##

-   **Basic Formula Syntax**

    -   $\sim$ and $+$: operators

    -   $y$: response variable

    -   $x, z$: explanatory variables

    -   $y \sim x + z$: a formula which means that $y$ depends on $x$ and $z$

        -   $+$ is used to indicate the addition of predictor variables to the right of the formula
        -   Be careful not to confuse the arithmetic operator $+$ with $+$ within a formula

\tiny

```{r}
#| eval: false
?`+` # Arithmetic Operators
?formula # operators in a formula
```

##

-   **Descriptives for n-Way Groups: the base R way**

    -   Split data into $n$ subsets and compute summary statistics

\tiny

```{r}
aggregate(x = income ~ Segment + ownHome, 
          data = segmentation, FUN = mean)
```

##

-   **Descriptives for n-Way Groups: the base R way**

    -   Split data into $n$ subsets and compute summary statistics

\tiny

```{r}
aggregate(x = kids ~ Segment + ownHome, 
          data = segmentation, FUN = sum)
```

##

-   **Descriptives for n-Way Groups: the tidyverse way**

    -   Split data into $n$ subsets and compute summary statistics

\tiny

```{r}
segmentation |> 
  group_by(Segment, ownHome) |>
  summarise(mean_income = mean(income))
```

##

-   **Descriptives for n-Way Groups: the tidyverse way**

    -   Split data into $n$ subsets and compute summary statistics

\tiny

```{r}
segmentation |> 
  group_by(Segment, ownHome) |>
  summarise(sum_kids = sum(kids))
```

##

-   **Basic Formula Syntax**

    -   $\sim$, $+$ and $|$: operators

    -   $y$: response variable

    -   $x$: explanatory variable

    -   $z$: grouping variable

    -   $y \sim x | z$: $y$ depends on $x$ based on different groups defined by $z$

        -   $|$ is used to separate the grouping variable from the explanatory variable
        -   Be careful not to confuse the logical operator $|$ with $|$ within a formula

\tiny

```{r}
#| eval: false
?`|` # Logical Operators
?lattice::xyplot # operators in a formula (you need first to install the package lattice)
```

##

-   **Visualization by group as frequencies: the lattice way**

\tiny

```{r}
library(lattice)
histogram(~ subscribe | Segment + ownHome, data = segmentation)
```

##

-   **Visualization by group as frequencies: the tidyverse way**

\tiny

```{r}
# Prepare data
subscriber_by_segment_home_ownership <- segmentation |>
  count(subscribe, Segment, ownHome) |>
  group_by(Segment, ownHome) |>
  mutate(n_pct = (n / sum(n)) * 100) |>
  ungroup()
subscriber_by_segment_home_ownership
```

##

-   **Visualization by group as frequencies: the tidyverse way**

\tiny

```{r}
subscriber_by_segment_home_ownership |>
  ggplot() + 
  geom_col(aes(x = subscribe, y=n_pct)) + 
  facet_wrap(facets = vars(Segment,ownHome), 
             nrow = 2, ncol = 4)
```

##

-   **Visualization by group as proportions: the lattice way**

\tiny

```{r}
# Prepare data
prop_table <- table(segmentation$subscribe,
                    segmentation$Segment) |>
  prop.table(margin = 2) |>
  _[2, ] # You can use _ as a placeholder. Check ?pipeOp
```

```{r}
barchart(prop_table, 
         xlab='Subscriber proportion by Segment', col='darkolivegreen')
```

##

-   **Visualization by group as proportions: the tidyverse way**

\tiny

```{r}
# Prepare data
prop_table <- segmentation |>
  count(subscribe, Segment) |>
  group_by(Segment) |>
  mutate(n_pct = n / sum(n)) |>
  filter(subscribe == 'subYes')
```

```{r}
prop_table |> ggplot() + 
  geom_col(aes(x=n_pct, y=Segment),
           fill='darkolivegreen') +
  labs(x='Subscriber proportion by Segment', 
       y=NULL)
```

##

-   **Visualization by group with continuous data: the lattice way**

\tiny

```{r}
# Prepare data
seg_income_agg <- aggregate(income ~ Segment + ownHome, 
                            data=segmentation, FUN = mean)
```

```{r}
#| echo: true
barchart(income ~ Segment, data = seg_income_agg,
         groups=ownHome, auto.key=TRUE, # Add groups
         par.settings=simpleTheme(col=terrain.colors(n = 2))) # Change default colors
```

##

-   **Visualization by group with continuous data: the tidyverse way**

\tiny

```{r}
# Prepare data
seg_income_agg <- segmentation |> 
  group_by(Segment, ownHome) |> 
  summarise(mean_income = mean(income)) |> 
  ungroup()
```

```{r}
seg_income_agg |> ggplot() +
  geom_col(aes(x=Segment, y=mean_income, fill=ownHome),
           position = position_dodge(), color='black') +
  scale_fill_manual(values=terrain.colors(n = 2))
```

##

-   **Visualization by group with continuous data: the lattice way**

\tiny

```{r}
bwplot(Segment ~ income | ownHome,
       data = segmentation,
       xlab = 'Income')
```

##

-   **Visualization by group with continuous data: the lattice way**

\tiny

```{r}
segmentation |> ggplot() +
  geom_boxplot(aes(x=income, y=Segment)) +
  facet_wrap(facets = vars(ownHome)) +
  labs(x='Income',
       y=NULL)
```

# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://posit.co/products/open-source/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

-   To the [**Linux kernel community**](https://www.kernel.org/category/about.html) for allowing me the possibility to use some [**Linux distributions**](https://static.lwn.net/Distributions/) as my main [**OS**](https://en.wikipedia.org/wiki/Operating_system) without paying for a license

# References {.allowframebreaks}
