---
title: "Reducing Data Complexity"
subtitle: Session 1
author:
  - name: 
      given: Luis F.
      family: Gómez L.
    orcid: 0000-0002-2877-9015
    email: luis.gomezl@unimilitar.edu.co
    affiliations:
      - name: Distance Learning Faculty
toc: true
toc-depth: 1
slide-level: 2
format:
  presentation-umng-revealjs:
    logo-vertical-strip: ../000_logos/logo-white-vertical.png 
    embed-resources: true
    mermaid-format: png
    reference-location: document
date: last-modified
date-format: "D MMMM, YYYY"
knitr: 
  opts_chunk: 
    echo: true
    warning: false
    fig.align: center
    out.width: 85%
lang: en
bibliography: ../000_references/marketing_research_and_analytics.bib
csl: ../000_references/apa-7.csl
editor: source
---

```{r}
#| echo: false
#| label: libraries

library(tidyverse)
library(skimr)
library(latex2exp)
library(ggforce)
library(gt)

library(kableExtra)
```

```{r}
#| label: palette
#| echo: false

umng_palette <- c(
  "#043074",
  "#fdc600",
  "#0d5e30",
  "#ee2a24",
  "#fc6700",
  "#00b3f0",
  "#6e3c1a",
  "#f8941c",
  "#8e44ad",
  "#2c3e50",
  "#16a085",
  "#c0392b",
  "#e91e63"
)
```

# Please Read Me

##

-   This presentation is based on [@chapman_r_2019, Chapter 11]

## 

- **Purpose**

    - Find groups of customers that differ in different dimensions to engage in more effective promotion

# Consumer segmentation survey

##



# Acknowledgments

##

-   To my family that supports me

-   To the taxpayers of Colombia and the [**UMNG students**](https://www.umng.edu.co/estudiante) who pay my salary

-   To the [**Business Science**](https://www.business-science.io/) and [**R4DS Online Learning**](https://www.rfordatasci.com/) communities where I learn [**R**](https://www.r-project.org/about.html) and [**$\pi$-thon**](https://www.python.org/about/)

-   To the [**R Core Team**](https://www.r-project.org/contributors.html), the creators of [**RStudio IDE**](https://posit.co/products/open-source/rstudio/), [**Quarto**](https://quarto.org/) and the authors and maintainers of the packages [**tidyverse**](https://CRAN.R-project.org/package=tidyverse), [**skimr**](https://CRAN.R-project.org/package=skimr), [**latex2exp**](https://CRAN.R-project.org/package=latex2exp), [**ggforce**](https://CRAN.R-project.org/package=ggforce),
[**gt**](https://CRAN.R-project.org/package=gt) and [**tinytex**](https://CRAN.R-project.org/package=tinytex) for allowing me to access these tools without paying for a license

# References {.allowframebreaks}

