---
title: "Marketing report part 1"
date: 2025-10-18
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gómez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| message: false

library(tidyverse)
library(sweep)
library(skimr)
```

# Describing data

## Introduction

Pending

## Import data

Pending

```{r}
# Import the data from the sweep package
bike_sales <- bike_sales
```

## Inspect data

Pending

```{r}
bike_sales |> 
  glimpse()
```

## Summarize data

Pending

### Summary statistics

Pending

```{r}
bike_sales |> 
  skim()
```

### Counting

Remember that you want to support the Research and Development (R & D) division in identifying potential new products

You can use the `count` function from the `tidyverse` and the variables `category.secondary` and `frame` to identify [product gaps](https://en.wikipedia.org/wiki/Gap_analysis). In that way you can identify possible new models that are not yet offered by the corporation but that can potentially generate an important revenue:

-   First, check the specific bicycle categories: `bike_sales |> count(category.secondary)`

-   Second, check the bicycle frame materials: `bike_sales |> count(frame)`

-   Third, you can calculate the possible combinations (Hint: There are $9 \times 2 = 18$ possible combinations)

Applying the above steps, you can proceed in the following way to identify [product gaps](https://en.wikipedia.org/wiki/Gap_analysis) (Hint: Only $13$ combinations are available which means that $18 - 13 = 5$ new models can be explored):

```{r}
#| code-fold: show
#| eval: false
# If you have loaded the tidyverse library, library(tidyverse),
## you can learn more about this function using ?count or
## check out https://dplyr.tidyverse.org/reference/count.html
## so see the documentation online
bike_sales |>
  count(category.secondary, frame)
```

Also remember that you want to support the marketing team in the creation of a marketing segmentation clustering model. Therefore the first step is to know the bicycle shops and the number of transactions. You can proceed in the following way as a first step:

```{r}
#| code-fold: show
#| eval: false

# Explore the argument sort in the count function
## You can check the documentation using ?count or
## checking it online 
### https://dplyr.tidyverse.org/reference/count.html
bike_sales |>
  count(bikeshop.name)
```

## Data visualization

Add some text pointing out the principal findings as a result of visualizing the suggested variables.

### Histograms

Check the distribution of the variables `quantity` and `price` to understand its behavior using the function `hist` from base R or the `tidyverse` through the package `ggplot2`

#### Quantity behavior

```{r}
# If you use base R check out the hist documentation
## ?hist
# If you use the tidyverse check out the geom_histogram
# documentation
## ?geom_histogram
### Also you can check out the online documentation in the 
### case of the tidyverse
#### https://ggplot2.tidyverse.org/reference/geom_histogram.html
```

#### Price behavior

```{r}
# Check out the comments pointed out in the 
# Quantity behavior code chunk section
```

### Boxplots

In this section generate 2 boxplots to explore the following items and point out the main finding in relation to these visualizations:

-   Examine the revenue generated by the corporation, `price.ext`, according to the specific bicycle categories, `category.secondary`

-   Examine the revenue generated by the corporation, `price.ext`, according to the bicycle frame material, `frame`

```{r}
# You can use the function boxplot, ?boxplot,from base R
# or the function geom_boxplot, ?geom_boxplot,from the 
# tidyverse
## If you use the tidyverse check out the 
## online documentation in
### https://ggplot2.tidyverse.org/reference/geom_boxplot.html
```

## Questions answered with data

This is a miscellaneous section where the idea is to answer questions using the `bike_sales` data set. Answer the following questions using the `tidyverse` and the functions `group_by`, `summarise`, `mutate`, `fct_reorder`, `ggplot` and `geom_col`:

-   What are the bike shops, `bikeshop.name`, that generate more revenue, `price.ext`, to the corporation?

-   What are the specific bicycle categories, `category.secondary`, that generate more revenue, `price.ext`, to the corporation?

To answer the above questions divide the process in 2 states:

-   First, prepare the data
-   Second, visualize the data

You can use the following example as a guidance where the variable `bikeshop.city` is used to answer what cities generate more revenue, `price.ext`:

```{r}
#| code-fold: show
#| eval: false

# Prepare data
bike_sales_total_revenue_by_city <- bike_sales |> 
  group_by(bikeshop.city) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  mutate(bikeshop.city = fct_reorder(.f = bikeshop.city,
                                     .x = total_revenue)) 

# Visualize data
bike_sales_total_revenue_by_city |> 
  ggplot(aes(x = total_revenue, y = bikeshop.city)) + 
  geom_col()
```
