---
title: "Marketing report part 1"
date: 2026-04-25
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gomez Lopez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| warning: false
#| message: false

library(tidyverse)
library(sweep)
library(skimr)
library(DT)
library(scales)
```

# Describing data

## Introduction

Pending bla bla bla

## Import data

Pending

```{r}
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## Inspect data

Pending

```{r}
bike_sales |> 
  glimpse()
```

## Summarize data

Pending

### Summary statistics

Pending

```{r}
bike_sales |> 
  skim()
```

### Counting

Pending

#### Potential products

Pending

```{r}
#| label: tbl-category-secondary-count
#| tbl-cap: Category secondary items
category_secondary_count <- bike_sales |> 
  count(category.secondary)

category_secondary_count |> 
  datatable(
    colnames = c(
      "Category Secondary",
      "Count"
    )
  )
```

```{r}
#| label: tbl-frame-count
#| tbl-cap: Frame items
frame_count <- bike_sales |> 
  count(frame)

frame_count |> 
  datatable(
    colnames = c(
      "Frame",
      "Count"
    )
  )
```

#### Products that are produced

```{r}
#| label: tbl-category-secondary-frame-count
#| tbl-cap: Combinations of Category Secondary and frames that are produced
category_secondary_frame_count <- bike_sales |> 
  count(
    category.secondary,
    frame
  )

category_secondary_frame_count  |> 
  datatable(
    colnames = c(
      "Category Secondary",
      "Frame",
      "Count"
    )
  )
```

#### New products

- Cyclocross Aluminum
- Fat Bike Carbon
- Over Mountain Aluminum
- Sport Carbon
- Triathalon Aluminum

#### Bike shops

```{r}
#| label: tbl-bikeshop-name-count
#| tbl-cap: Bikeshops
bikeshop_name_count <- bike_sales |> 
  count(bikeshop.name)

bikeshop_name_count  |> 
  datatable(
    colnames = c(
      "Bike shop",
      "Count"
    )
  )
```

## Data visualization

Pending

### Histograms

Pending

#### Quantity behavior

Pending

```{r}
#| message: false
bike_sales |> 
  ggplot() +
  geom_histogram(
    aes(x = quantity),
    color = "black"
  ) + 
  labs(
    x = "Sales (units)",
    y = "Frequency",
    subtitle = "Distribution of sales in units of frames"
  )
```

#### Price behavior

Pending

```{r}
#| message: false
bike_sales |> 
  ggplot() +
  geom_histogram(
    aes(x = price),
    color = "black"
  ) +
  scale_x_continuous(
    labels = label_currency()
  ) +
  labs(
    x = "Frame prices (US Dollars)",
    y = "Frequency",
    subtitle = "Distribution of frame prices"
  )
```

### Boxplots

Pending

#### Revenue vs Category secondary

Pending

```{r}
bike_sales |> 
  ggplot() +
  geom_boxplot(
    aes(x = price.ext, y = category.secondary)
  ) +
  scale_x_continuous(
    labels = label_currency(),
    transform = "log10"
  ) +
  labs(
    x = "Revenue (US Dollars)",
    y = "Category Secondary",
    subtitle = "Distribution of revenue by category secondary"
  )
```

#### Revenue vs frame

Pending

```{r}
bike_sales |> 
  ggplot() +
  geom_boxplot(
    aes(x = price.ext, y = frame)
  ) +
  scale_x_continuous(
    label = label_currency(),
    transform = "log10"
  ) +
  labs(
    x = "Revenue (US Dollars)",
    y = "Frame",
    subtitle = "Distribution of revenue by frame"
  )
```

## Questions answered with data

Pending

-   What are the bike shops, `bikeshop.name`, that generate more revenue, `price.ext`, to the corporation?

Pending

```{r}
total_revenue_by_bikeshops <- bike_sales |> 
  group_by(bikeshop.name) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  mutate(bikeshop.name = fct_reorder(
    .f = bikeshop.name, .x = total_revenue
  )
)

total_revenue_by_bikeshops |> 
  ggplot() +
  geom_col(
    aes(x = total_revenue, y = bikeshop.name),
    color = "black"
  ) +
  scale_x_continuous(
    label = label_currency()
  ) +
  labs(
    x = "Total revenue (US Dollars)",
    y = "Bike Shop",
    subtitle = "Revenue by bike shop"
  )
```

-   What are the specific bicycle categories, `category.secondary`, that generate more revenue, `price.ext`, to the corporation?

Pending

```{r}
total_revenue_by_category_secondary <- bike_sales |> 
  group_by(category.secondary) |> 
  summarise(total_revenue = sum(price.ext)) |> 
  mutate(category.secondary = fct_reorder(
    .f = category.secondary, .x = total_revenue
  )
)

total_revenue_by_category_secondary |> 
  ggplot() +
  geom_col(
    aes(x = total_revenue, y = category.secondary),
    color = "black"
  ) +
  scale_x_continuous(
    label = label_currency()
  ) +
  labs(
    x = "Total revenue (US Dollars)",
    y = "Category secondary",
    subtitle = "Revenue by category secondary"
  )
```

# Ai use disclaimer

Our group used the following artificial intelligence tools:

- Gemini 
- Chatgpt
- Claude

Our group apply the following uses:

- [x] Assistance in text writing
- [ ] Assistance in image creation
- [x] Assistance in developing ideas and exploring arguments
- [ ] Assistance with computer programming