---
title: "Marketing report part 1"
date: 2026-04-25
author: 
  - Norah Jones   
  - Bill Gates
  - Luis Francisco Gomez Lopez
format: 
  html:
    toc: true
    code-fold: true
    number-sections: true
    embed-resources: true
editor: source
---

```{r }
#| label: libraries
#| warning: false
#| message: false

library(tidyverse)
library(sweep)
library(skimr)
library(DT)
```

# Describing data

## Introduction

Pending bla bla bla

## Import data

Pending

```{r}
bike_sales <- bike_sales |> 
  mutate(
    bikeshop.name      = factor(x = bikeshop.name, ordered = FALSE),
    bikeshop.city      = factor(x = bikeshop.city, ordered = FALSE),
    bikeshop.state     = factor(x = bikeshop.state, ordered = FALSE),
    model              = factor(x = model, ordered = FALSE),
    category.primary   = factor(x = category.primary, ordered = FALSE),
    category.secondary = factor(x = category.secondary, ordered = FALSE),
    frame              = factor(x = frame, ordered = FALSE),
    order.id           = factor(x = order.id, ordered = FALSE),
    order.line         = factor(x = order.line, ordered = TRUE),
    customer.id        = factor(x = customer.id, ordered = FALSE),
    product.id         = factor(x = product.id, ordered = FALSE)
  )
```

## Inspect data

Pending

```{r}
bike_sales |> 
  glimpse()
```

## Summarize data

Pending

### Summary statistics

Pending

```{r}
bike_sales |> 
  skim()
```

### Counting

Pending

#### Potential products

Pending

```{r}
#| label: tbl-category-secondary-count
#| tbl-cap: Category secondary items
category_secondary_count <- bike_sales |> 
  count(category.secondary)

category_secondary_count |> 
  datatable(
    colnames = c(
      "Category Secondary",
      "Count"
    )
  )
```

```{r}
#| label: tbl-frame-count
#| tbl-cap: Frame items
frame_count <- bike_sales |> 
  count(frame)

frame_count |> 
  datatable(
    colnames = c(
      "Frame",
      "Count"
    )
  )
```

#### Products that are produced

```{r}
#| label: tbl-category-secondary-frame-count
#| tbl-cap: Combinations of Category Secondary and frames that are produced
category_secondary_frame_count <- bike_sales |> 
  count(
    category.secondary,
    frame
  )

category_secondary_frame_count  |> 
  datatable(
    colnames = c(
      "Category Secondary",
      "Frame",
      "Count"
    )
  )
```

#### New products

- Cyclocross Aluminum
- Fat Bike Carbon
- Over Mountain Aluminum
- Sport Carbon
- Triathalon Aluminum

#### Bike shops

```{r}
#| label: tbl-bikeshop-name-count
#| tbl-cap: Bikeshops
bikeshop_name_count <- bike_sales |> 
  count(bikeshop.name)

bikeshop_name_count  |> 
  datatable(
    colnames = c(
      "Bike shop",
      "Count"
    )
  )
```

## Data visualization

Add some text pointing out the principal findings as a result of visualizing the suggested variables.

### Histograms

Check the distribution of the variables `quantity` and `price` to understand its behavior using the function `hist` from base R or the `tidyverse` through the package `ggplot2`

#### Quantity behavior

```{r}
# If you use base R check out the hist documentation
## ?hist
# If you use the tidyverse check out the geom_histogram
# documentation
## ?geom_histogram
### Also you can check out the online documentation in the
### case of the tidyverse
#### https://ggplot2.tidyverse.org/reference/geom_histogram.html
```

#### Price behavior

```{r}
# Check out the comments pointed out in the
# Quantity behavior code chunk section
```

### Boxplots

In this section generate 2 boxplots to explore the following items and point out the main finding in relation to these visualizations:

-   Examine the revenue generated by the corporation, `price.ext`, according to the specific bicycle categories, `category.secondary`

-   Examine the revenue generated by the corporation, `price.ext`, according to the bicycle frame material, `frame`

```{r}
# You can use the function boxplot, ?boxplot,from base R
# or the function geom_boxplot, ?geom_boxplot,from the
# tidyverse
## If you use the tidyverse check out the
## online documentation in
### https://ggplot2.tidyverse.org/reference/geom_boxplot.html
```

## Questions answered with data

This is a miscellaneous section where the idea is to answer questions using the `bike_sales` data set. Answer the following questions using the `tidyverse` and the functions `group_by`, `summarise`, `mutate`, `fct_reorder`, `ggplot` and `geom_col`:

-   What are the bike shops, `bikeshop.name`, that generate more revenue, `price.ext`, to the corporation?

-   What are the specific bicycle categories, `category.secondary`, that generate more revenue, `price.ext`, to the corporation?

To answer the above questions divide the process in 2 states:

-   First, prepare the data
-   Second, visualize the data

You can use the following example as a guidance where the variable `bikeshop.city` is used to answer what cities generate more revenue, `price.ext`:

```{r}
#| code-fold: show
#| eval: false

# Prepare data
bike_sales_total_revenue_by_city <- bike_sales |>
  group_by(bikeshop.city) |>
  summarise(total_revenue = sum(price.ext)) |>
  mutate(bikeshop.city = fct_reorder(.f = bikeshop.city, .x = total_revenue))

# Visualize data
bike_sales_total_revenue_by_city |>
  ggplot(aes(x = total_revenue, y = bikeshop.city)) +
  geom_col()
```

# Ai use disclaimer

In this section specify in a clear and explicit way if any artificial intelligence tools where used. 

In case any artificial intelligency tools where used point out how they were use according to the following authorized uses (uncheck the uses that doesn't apply):

- [x] Assistance in text writing
- [x] Assistance in image creation
- [x] Assistance in developing ideas and exploring arguments
- [x] Assistance with computer programming